---
title: "S1 Chlorophyll-a bloom detection"
unnumbered: true
---

```{r}
#| label: setup-bloom-onset
#| include: false

# For data manipulation and plotting (loads ggplot2, dplyr, tidyr, etc.)
library(tidyverse)

# For handling and plotting spatial data
library(sf)
library(rnaturalearth)
library(ggspatial)

# For combining plots
library(cowplot)   # Used for the map insets
library(patchwork) # Used for the final density plots

# For file path management
library(here)

# For time-series calculations (rolling medians)
library(zoo)

# For creating interactive HTML tables
library(DT)


```

# S1.1 Map of Isla Isabel

A map of Isla Isabel, Mexico, is provided for geographic context. The figure shows the island’s location in the eastern tropical Pacific, the surrounding coastal region, and the focal extraction area used for chlorophyll-a data.

```{r}
#| label: map_final_refinements
#| fig-width: 10
#| fig-height: 8
#| code-fold: true



# --- Define Data ---
# Define the polygon coordinates
latitudes <- c(22.1203, 22.1203, 21.5797, 21.5797, 22.1203)
longitudes <- c(-106.1763, -105.5937, -105.5937, -106.1763, -106.1763)
coords_df <- data.frame(lon = longitudes, lat = latitudes)
polygon_sf <- st_polygon(list(as.matrix(coords_df)))
polygon_sf <- st_sfc(polygon_sf, crs = 4326)

# Define the marker coordinates for Isla Isabel
marker_lat_deg <- 21 + 50/60 + 59.0/3600
marker_lon_deg <- -(105 + 52/60 + 54.0/3600)
marker_point_df <- data.frame(lon = marker_lon_deg, lat = marker_lat_deg)
marker_point_sf <- st_point(c(marker_point_df$lon, marker_point_df$lat))
marker_point_sf <- st_sfc(marker_point_sf, crs = 4326)

# Get world map data
world <- ne_countries(scale = "medium", returnclass = "sf")
coastlines <- ne_coastline(scale = "medium", returnclass = "sf")

# Create points for country labels
world_points <- cbind(world, st_coordinates(st_centroid(world$geometry)))

# --- Main Map (Zoomed Out - Pacific Coast) ---
main_map <- ggplot() +
  # TYPO FIX: Changed "antiquhite" to "antiquewhite"
  geom_sf(data = world, fill = "antiquewhite", color = "darkgray") +
  geom_sf(data = coastlines, color = "grey", size = 0.5) +
  geom_sf(data = st_as_sfc(st_bbox(polygon_sf)), fill = NA, color = "red", size = 1) +
  geom_text(data = world_points, aes(x = X, y = Y, label = name),
            color = "black", size = 3.5, fontface = "bold", check_overlap = TRUE) +
  annotate(geom = "text", x = -90, y = 25, label = "Gulf of Mexico",
           fontface = "italic", color = "steelblue4", size = 5) +
  coord_sf(xlim = c(-125, -85), ylim = c(12, 35), expand = FALSE) +
  labs(title = "", x = "Longitude (°W)", y = "Latitude (°N)") +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    panel.grid.major = element_line(color = "lightgray", linetype = "dashed"),
    panel.background = element_rect(fill = "aliceblue")
  )

# --- Inset Map (Zoomed In on Isla Isabel) ---
inset_map <- ggplot() +
  geom_sf(data = world, fill = "antiquewhite", color = "darkgray") +
  geom_sf(data = coastlines, color = "grey", size = 0.5) +
  geom_sf(data = polygon_sf, fill = "blue", alpha = 0.5, color = "darkblue", size = 1) +
  geom_sf(data = marker_point_sf, color = "red", size = 4, shape = 17) +
  geom_text(data = marker_point_df, aes(x = lon, y = lat), label = "Isla Isabel", hjust = 0.5, vjust = -5.5, color = "black", fontface = "bold", size = 6) +
  annotation_scale(location = "bl", width_hint = 0.2) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.09, "in"), pad_y = unit(0.3, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(
    xlim = c(marker_lon_deg - 0.5, marker_lon_deg + 0.5),
    ylim = c(marker_lat_deg - 0.5, marker_lat_deg + 0.5),
    expand = FALSE
  ) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "aliceblue"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
  )

# --- Combine Maps using cowplot ---
mapIS<-ggdraw() +
  draw_plot(main_map) +
  draw_plot(inset_map, x = 0.05, y = 0.15, width = 0.38, height = 0.38)

```



::: card
```{r}
mapIS
```
:::

```{r}
#| label: save_map
#| eval: false

ggsave(filename = here("..","Plots", "Map_Isla_Isabel.pdf"),
       plot = mapIS,
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "Map_Isla_Isabel.jpg"),
       plot = mapIS,
       width = 7,  
       height = 7,  
       dpi = 300)
```

# S1.2 Bloom onset
This section describes the procedure used to estimate phytoplankton bloom onset from satellite-derived chlorophyll-a time series.

## Chlorophyll-a bloom detection

Satellite-derived chlorophyll-a data were processed to identify the onset of seasonal blooms.

::::: panel-tabset
## Load chlorophyll data

Chlorophyll-a data were loaded from a CSV file, cleaned by removing missing values (NA), restricted to the date and chlorophyll columns, and converted to a date format using `lubridate::dmy()`.


```{r}
#| label: chlorophyll data
CHL_df <- read.csv(here("..","data", "MODIS.CHL.R2022.csv"), header = TRUE) %>%
  drop_na() %>%
  dplyr::select(date, chlorophyll) %>%
  rename(daily_value = chlorophyll)

CHL_df <- CHL_df %>%
  mutate(date = dmy(date)) 
```

## Estimate sustained bloom onset

Two custom functions were used to estimate bloom onset dates under different seasonal-window definitions. Both apply the same detection procedure:

- Seasonal window: Define a season year spanning either Winter–Spring (21 December–20 June) or Autumn–Spring (21 September–20 June).

- Baseline: Compute the seasonal median chlorophyll-a concentration for each season year.

- Thresholds: Define dynamic thresholds at 5%, 15%, 25%, and 30% above the seasonal median.

- Sustained increase criterion: Identify the first day for which two consecutive 7-day rolling medians exceed the threshold, to distinguish sustained increases from short-term fluctuations.

- Output: Return the estimated bloom onset date, chlorophyll-a value at onset, and the corresponding threshold value.

:::: panel-tabset
Winter–Spring window (21 December–20 June)

This function defines the season as running from December 21st of the previous year to June 20th of the current year.

```{r}
#| label: Function to calculate sustained bloom start date

calculate_sustained_bloom_start <- function(data_frame, value_column_name = "daily_value", min_duration_days = 7) {

  thresholds <- c(0.05, 0.15,0.25, 0.30)

  # Prepare and annotate data
  data_frame <- data_frame %>%
    mutate(date = as.Date(date)) %>%
    arrange(date) %>%
    mutate(season_year = ifelse(month(date) == 12, year(date) + 1, year(date)))

  # Define seasonal window
  season_ranges <- data_frame %>%
    distinct(season_year) %>%
    mutate(
      season_start_date = ymd(paste0(season_year - 1, "-12-21")),
      season_end_date = ymd(paste0(season_year, "-06-20"))
    )

  # Filter seasonal data
  seasonal_data <- data_frame %>%
    left_join(season_ranges, by = "season_year") %>%
    filter(date >= season_start_date & date <= season_end_date) %>%
    arrange(season_year, date)

  # Compute yearly medians
  medians <- seasonal_data %>%
    group_by(season_year) %>%
    summarise(yearly_median = median(!!sym(value_column_name), na.rm = TRUE), .groups = "drop")

  # Initialize final result with medians
  all_results <- medians
  bloom_date_raws <- list()

  for (thr in thresholds) {
    label <- paste0("thr", thr * 100)

    # Add threshold column
    threshold_tbl <- medians %>%
      mutate(threshold = yearly_median * (1 + thr))

    seasonal_thr_data <- seasonal_data %>%
      left_join(threshold_tbl, by = "season_year")

    # Compute rolling medians
    seasonal_thr_data <- seasonal_thr_data %>%
      group_by(season_year) %>%
      mutate(
        rolling_median_week1 = rollapply(
          !!sym(value_column_name), min_duration_days, median, na.rm = TRUE, fill = NA, align = "left"
        ),
        rolling_median_week2 = rollapply(
          lead(!!sym(value_column_name), n = min_duration_days), min_duration_days, median, na.rm = TRUE, fill = NA, align = "left"
        ),
        cond_i = !is.na(rolling_median_week1) & rolling_median_week1 >= threshold,
        cond_ii = !is.na(rolling_median_week2) & rolling_median_week2 >= threshold
      ) %>%
      filter(cond_i & cond_ii) %>%
      summarise(bloom_date = min(date, na.rm = TRUE), .groups = "drop")

    # Store raw date for difference calculation
    bloom_date_raws[[label]] <- seasonal_thr_data %>%
      rename(!!paste0("bloom_date_raw_", label) := bloom_date)

    # Join back to get bloom values and threshold
    enriched <- seasonal_thr_data %>%
      left_join(seasonal_data, by = "season_year") %>%
      group_by(season_year) %>%
      summarise(
        !!paste0("bloom_start_", label) := format(first(bloom_date), "%d/%m/%Y"),
        !!paste0("value_at_bloom_", label) := seasonal_data[[value_column_name]][match(first(bloom_date), seasonal_data$date)],
        !!paste0("threshold_", label) := medians$yearly_median[medians$season_year == first(season_year)] * (1 + thr),
        .groups = "drop"
      )

    # Join enriched info to results
    all_results <- all_results %>%
      left_join(enriched, by = "season_year")
  }

  # Join all bloom date raws
  bloom_date_df <- reduce(bloom_date_raws, left_join, by = "season_year")

  # Add differences between 5% vs 15% and 30%
  all_results <- all_results %>%
    left_join(bloom_date_df, by = "season_year") %>%
    mutate(
      diff_days_5_vs_15 = as.numeric(bloom_date_raw_thr15 - bloom_date_raw_thr5),
      diff_days_5_vs_30 = as.numeric(bloom_date_raw_thr30 - bloom_date_raw_thr5)
    ) %>%
    select(-starts_with("bloom_date_raw_"))

  return(all_results)
}
 bloom_startsA <- calculate_sustained_bloom_start(CHL_df)
#write.csv(bloom_starts2, here("data", "bloom_dates.csv"), row.names = FALSE)
```

## Autumn–Spring window (21 September–20 June)

This function defines a broader season, running from September 21st of the previous year to June 20th of the current year.

```{r}
#| label: Function to calculate sustained bloom start from Autumn to Spring


calculate_sustained_bloom_startB <- function(data_frame, value_column_name = "daily_value", min_duration_days = 7) {

  thresholds <- c(0.05, 0.15,0.25, 0.30)

  # Prepare and annotate data
  data_frame <- data_frame %>%
    mutate(date = as.Date(date)) %>%
    arrange(date) %>%
    mutate(season_year = ifelse(month(date) == 9, year(date) + 1, year(date)))

  # Define seasonal window
  season_ranges <- data_frame %>%
    distinct(season_year) %>%
    mutate(
      season_start_date = ymd(paste0(season_year - 1, "-09-21")),
      season_end_date = ymd(paste0(season_year, "-06-20"))
    )

  # Filter seasonal data
  seasonal_data <- data_frame %>%
    left_join(season_ranges, by = "season_year") %>%
    filter(date >= season_start_date & date <= season_end_date) %>%
    arrange(season_year, date)

  # Compute yearly medians
  medians <- seasonal_data %>%
    group_by(season_year) %>%
    summarise(yearly_median = median(!!sym(value_column_name), na.rm = TRUE), .groups = "drop")

  # Initialize final result with medians
  all_results <- medians
  bloom_date_raws <- list()

  for (thr in thresholds) {
    label <- paste0("thr", thr * 100)

    # Add threshold column
    threshold_tbl <- medians %>%
      mutate(threshold = yearly_median * (1 + thr))

    seasonal_thr_data <- seasonal_data %>%
      left_join(threshold_tbl, by = "season_year")

    # Compute rolling medians
    seasonal_thr_data <- seasonal_thr_data %>%
      group_by(season_year) %>%
      mutate(
        rolling_median_week1 = rollapply(
          !!sym(value_column_name), min_duration_days, median, na.rm = TRUE, fill = NA, align = "left"
        ),
        rolling_median_week2 = rollapply(
          lead(!!sym(value_column_name), n = min_duration_days), min_duration_days, median, na.rm = TRUE, fill = NA, align = "left"
        ),
        cond_i = !is.na(rolling_median_week1) & rolling_median_week1 >= threshold,
        cond_ii = !is.na(rolling_median_week2) & rolling_median_week2 >= threshold
      ) %>%
      filter(cond_i & cond_ii) %>%
      summarise(bloom_date = min(date, na.rm = TRUE), .groups = "drop")

    # Store raw date for difference calculation
    bloom_date_raws[[label]] <- seasonal_thr_data %>%
      rename(!!paste0("bloom_date_raw_", label) := bloom_date)

    # Join back to get bloom values and threshold
    enriched <- seasonal_thr_data %>%
      left_join(seasonal_data, by = "season_year") %>%
      group_by(season_year) %>%
      summarise(
        !!paste0("bloom_start_", label) := format(first(bloom_date), "%d/%m/%Y"),
        !!paste0("value_at_bloom_", label) := seasonal_data[[value_column_name]][match(first(bloom_date), seasonal_data$date)],
        !!paste0("threshold_", label) := medians$yearly_median[medians$season_year == first(season_year)] * (1 + thr),
        .groups = "drop"
      )

    # Join enriched info to results
    all_results <- all_results %>%
      left_join(enriched, by = "season_year")
  }

  # Join all bloom date raws
  bloom_date_df <- reduce(bloom_date_raws, left_join, by = "season_year")

  # Add differences between 5% vs 15% and 30%
  all_results <- all_results %>%
    left_join(bloom_date_df, by = "season_year") %>%
    mutate(
      diff_days_5_vs_15 = as.numeric(bloom_date_raw_thr15 - bloom_date_raw_thr5),
      diff_days_5_vs_30 = as.numeric(bloom_date_raw_thr30 - bloom_date_raw_thr5)
    ) %>%
    select(-starts_with("bloom_date_raw_"))

  return(all_results)
}
bloom_startsB <- calculate_sustained_bloom_startB(CHL_df)
#write.csv(bloom_startsB, here("data", "bloom_datesB.csv"), row.names = FALSE)
```

## Summary of bloom start dates

::: panel-tabset
## Summary for Winter to Spring
This table summarizes estimated bloom onset dates for the Winter–Spring seasonal window (21 December–20 June).

```{r}
#| label: Summary A
#| echo: false
DT::datatable(
  bloom_startsA,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv')
  ),
  extensions = 'Buttons',
  rownames = FALSE
) %>%
  formatRound(columns = c("yearly_median","value_at_bloom_thr5", "threshold_thr5","value_at_bloom_thr15","threshold_thr15","value_at_bloom_thr25","threshold_thr25","value_at_bloom_thr30","threshold_thr30"), digits = 3)%>%
  formatStyle(columns = names(bloom_startsA), `text-align` = 'center')
```

## Summary for Autumn to Spring

This table summarizes estimated bloom onset dates for the Autumn–Spring seasonal window (21 September–20 June).

```{r}
#| label: Summary B
#| echo: false
DT::datatable(
  bloom_startsB,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv')
  ),
  extensions = 'Buttons',
  rownames = FALSE
) %>%
  formatRound(columns = c("yearly_median","value_at_bloom_thr5", "threshold_thr5","value_at_bloom_thr15","threshold_thr15","value_at_bloom_thr25","threshold_thr25","value_at_bloom_thr30","threshold_thr30"), digits = 3)%>%
  formatStyle(columns = names(bloom_startsB), `text-align` = 'center')
```
:::
::::
:::::

# S1.3 Visualization of Chlorophyl-a bloom onset dates

To visualize the distribution of estimated bloom onset dates across years, detection outputs were transformed and plotted for comparison across thresholds and seasonal windows.

### Data processing for visualization

Before plotting, bloom onset estimates were converted to long format for visualization with `ggplot2`. A helper function (`process_for_plotting`) standardized dates to a common seasonal calendar by assigning September–December dates to year 2000 and January–June dates to year 2001, enabling comparison of timing distributions across years and thresholds.

```{r}
#| label: Frequency of bloom start dates
#| code-fold: true

process_for_plotting <- function(results) {
  results %>%
    select(season_year, starts_with("bloom_start_")) %>%
    pivot_longer(
      cols = -season_year,
      names_to = "threshold_id",
      values_to = "bloom_date_str",
      values_drop_na = TRUE
    ) %>%
    mutate(
      threshold_label = gsub("bloom_start_thr", "", threshold_id),
      threshold = factor(paste0(threshold_label, "%"), levels = c("5%", "15%", "25%", "30%")),
      bloom_date = dmy(bloom_date_str),
      bloom_day_month = if_else(
          month(bloom_date) >= 9,
          update(bloom_date, year = 2000),
          update(bloom_date, year = 2001)
      )
    )
}

bloom_dates_long_A <- process_for_plotting(bloom_startsA)
bloom_dates_long_B <- process_for_plotting(bloom_startsB)


sample_sizes_A <- bloom_dates_long_A %>%
  group_by(threshold) %>%
  summarise(label = paste("n =", n_distinct(season_year)), .groups = "drop")

sample_sizes_B <- bloom_dates_long_B %>%
  group_by(threshold) %>%
  summarise(label = paste("n =", n_distinct(season_year)), .groups = "drop")


mean_dates_A <- bloom_dates_long_A %>%
  group_by(threshold) %>%
  summarise(mean_date = as.Date(mean(as.numeric(bloom_day_month)), origin = "1970-01-01"), .groups = "drop")

mean_dates_B <- bloom_dates_long_B %>%
  group_by(threshold) %>%
  summarise(mean_date = as.Date(mean(as.numeric(bloom_day_month)), origin = "1970-01-01"), .groups = "drop")



```

### Generating and combining onset-date distribution plots

Binned area plots were generated for each seasonal window and faceted by threshold level. Panel labels indicate the number of seasons with a detected bloom onset for each threshold. The Winter–Spring and Autumn–Spring panels were combined for side-by-side comparison.
```{r}
#| label: Area plot of bloom start dates

# --- Define the color palette once ---
bloom_colors <- c("5%"="#2c7bb6", "15%"="#abd9e9", "25%"="#fdae61", "30%"="#d7191c")

# --- Plot A ---
area_plot_A <- ggplot(bloom_dates_long_A, 
                      # Add 'color' aesthetic
                      aes(x = bloom_day_month, fill = threshold, color = threshold)) +
  

  geom_area(
    stat = "bin",
    binwidth = 1,         
    boundary = 0,
    alpha = 0.6,          # Adjusted alpha for better overlap
    position = "identity",
    linewidth = 0.5       # Controls outline thickness
  ) +
  
  geom_text(data = sample_sizes_A, aes(x = as.Date("2001-06-15"), y = Inf, label = label), vjust = 1.5, hjust = 1, size = 3, inherit.aes = FALSE) +
  facet_wrap(~threshold, ncol = 1, scales = "fixed") +
  labs(subtitle = "Winter to Spring", x = NULL, y = "Count") +
  theme_light() +
  scale_fill_manual(values = bloom_colors) +
  scale_color_manual(values = bloom_colors) + 
  scale_x_date(limits = c(as.Date("2000-09-01"), as.Date("2001-06-20")), date_labels = "%b", date_breaks = "1 month") +
  theme(plot.subtitle = element_text(hjust = 0.5), legend.position = "none", strip.text = element_text(face = "bold"))

# --- Plot B ---
area_plot_B <- ggplot(bloom_dates_long_B, 
                    
                      aes(x = bloom_day_month, fill = threshold, color = threshold)) +
  

  geom_area(
    stat = "bin",
    binwidth = 1,         
    boundary = 0,
    alpha = 0.6,
    position = "identity",
    linewidth = 0.5
  ) +
  
  facet_wrap(~threshold, ncol = 1, scales = "fixed") + 
  geom_text(data = sample_sizes_B, aes(x = as.Date("2001-06-15"), y = Inf, label = label), vjust = 1.5, hjust = 1, size = 3, inherit.aes = FALSE)+
  labs(subtitle = "Autumn to Spring", x = NULL, y = "Count") +
  theme_light() +
  scale_fill_manual(values = bloom_colors) +
  scale_color_manual(values = bloom_colors) + 
  scale_x_date(limits = c(as.Date("2000-09-01"), as.Date("2001-06-20")), date_labels = "%b", date_breaks = "1 month") +
  theme(plot.subtitle = element_text(hjust = 0.5), legend.position = "none", strip.text = element_text(face = "bold"))

# --- Combine plots ---
combined_plot <- area_plot_A + area_plot_B +
  plot_annotation(
    tag_levels = 'A'
  )


```

::: card
```{r}
# Print the final combined plot
print(combined_plot)
```
:::





### Description of bloom onset date distributions

This figure summarizes the distribution of detected phytoplankton bloom onset dates across years under alternative detection thresholds and seasonal windows. The left column shows results for the Winter–Spring window (21 December–20 June), and the right column shows results for the Autumn–Spring window (21 September–20 June). Rows correspond to threshold definitions (5%, 15%, 25%, and 30% above the seasonal median chlorophyll-a concentration).

Bloom onset dates were standardized to a common seasonal calendar (September–December mapped to year 2000; January–June mapped to year 2001) to enable comparison of timing distributions across years. Panel labels report the number of seasons with detected bloom onsets for each threshold.

The plots are provided to visualize how estimated bloom onset timing varies across threshold choices and seasonal-window definitions, and to support the robustness comparisons used in the main analyses and supplementary sensitivity checks.



```{r}
#| label: Save bloom start date plots
#| eval: false
ggsave(filename = here("..","Plots", "bloom_thresholds.pdf"),
       plot = combined_plot,
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "bloom_thresholds.jpg"),
       plot = combined_plot,
       width = 7,  
       height = 7,  
       dpi = 300)
```

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
