---
title: "S2. Phenological mismatch"
---

```{r}
#| label: setup-pheno-mismatch
#| include: false

# For data manipulation and plotting (loads ggplot2, dplyr, tidyr, etc.)
library(tidyverse)

# For file path management
library(here)
```
This section describes the preparation and visualization of phenological mismatch values, defined as the difference (in days) between laying date and estimated phytoplankton bloom onset.

## Data preparation

Phenological mismatch values were compiled for females and males using alternative bloom-onset definitions (5%, 15%, and 25% thresholds above the seasonal median chlorophyll-a concentration) and two seasonal windows (Winter–Spring and Autumn–Spring).

For visualization, the dataset was reshaped to long format for plotting in `ggplot2`, producing a combined dataset with mismatch values grouped by sex, seasonal window, and threshold definition.

## Mismatch visualization

Phenological mismatches were visualized as density distributions. Each panel corresponds to a combination of sex (females or males) and seasonal window (Winter–Spring or Autumn–Spring), with distributions compared across thresholds (5%, 15%, and 25%).

```{r}
#| label: Plot pheno mismatch
#| code-fold: true
dat <- read.csv(here::here("..","data", "Lay_date_mismatches_250702.csv"))


# 1. Prepare 'Winter to Spring' data
data_A <- dat %>%
  select(SEX, X5MISMATCH, X15MISMATCH, X25MISMATCH) %>%
  rename(
    `5% threshold` = `X5MISMATCH`,
    `15% threshold` = `X15MISMATCH`,
    `25% threshold` = `X25MISMATCH`
  ) %>%
  pivot_longer(
    cols = -SEX,
    names_to = "Mismatch_Type",
    values_to = "Value"
  ) %>%
  mutate(Season = "Winter to Spring")

# 2. Prepare 'Autumn to Spring' data
data_B <- dat %>%
  select(SEX, X5MISMATCHB, X15MISMATCHB) %>%
  rename(
    `5% threshold` = `X5MISMATCHB`,
    `15% threshold` = `X15MISMATCHB`
  ) %>%
  pivot_longer(
    cols = -SEX,
    names_to = "Mismatch_Type",
    values_to = "Value"
  ) %>%
  mutate(Season = "Autumn to Spring")


combined_data <- bind_rows(data_A, data_B) %>%
  mutate(

    Sex_Label = factor(if_else(SEX == 0, "Males", "Females"), levels = c("Females", "Males")),

    Season = factor(Season, levels = c("Winter to Spring", "Autumn to Spring")),

    Mismatch_Type = factor(Mismatch_Type, levels = c("5% threshold", "15% threshold", "25% threshold"))
  ) %>%

  filter(!is.na(Value))



```

```{r}
# --- NEW 2-COLOR PALETTE ---
season_palette <- c(
  "Winter to Spring" = "#0072B2", # Blue
  "Autumn to Spring" = "#DAA520"  # Gold (Goldenrod)
)

# --- MODIFIED PLOT CODE ---

# Map fill and color to 'Season'
mismatch_plot <- ggplot(combined_data, aes(x = Value, 
                                          fill = Season, 
                                          color = Season)) +
  
  geom_area(
    stat = "bin",
    binwidth = 1,
    position = "identity",
    alpha = 0.5, # Alpha is critical for seeing the overlap
    linewidth = 0
  ) +
  
  # Faceting remains the same (3 rows, 2 cols)
 facet_grid(Mismatch_Type ~ Sex_Label) +
  
 labs(
  title = "",
  x = "Phenological mismatch (days)",
  y = "Count"
 ) +
  
  # --- Use the new 2-color palette ---
 scale_fill_manual(
    name = "Season", 
    values = season_palette, 
    drop = FALSE
  ) +
  scale_color_manual(
    name = "Season", 
    values = season_palette, 
    drop = FALSE
  ) +
  
 theme_light() +
 
 scale_x_continuous(limits = c(-150, 250), breaks = seq(-150, 250, 50)) +
 
  theme(
  legend.position = "right",
    strip.text = element_text(face = "bold")
  )


```

::: card
```{r}
print(mismatch_plot)
```
:::


```{r}
ggsave(filename = here("..","Plots", "mismatch_thresholds.pdf"),
       plot = mismatch_plot,
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "mismatch_thresholds.jpg"),
       plot = mismatch_plot,
       width = 9,  
       height = 7,  
       dpi = 300)
```

## Distributional comparison of mismatch values

These density-style histograms compare the distribution of phenological mismatch values across bloom-onset thresholds (5%, 15%, 25%), sexes, and seasonal windows (Winter–Spring vs Autumn–Spring). Panels are faceted by threshold (rows) and sex (columns), and bars are colored by seasonal window. This figure is used to compare how mismatch distributions vary under alternative bloom definitions used in the sensitivity analyses.

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
