---
title: "S3. Model fitting workflow"
---

This section documents the model-fitting workflow used in the study. It includes:

-   Location–scale models for bloom phenology (Gaussian outcomes)
-   Hierarchical Bayesian models for phenological mismatch (Gaussian) and reproductive performance (ordinal cumulative probit)

Most code chunks are provided as a reproducible workflow record (`eval: false`) rather than executed during rendering of the supplementary document.

```{r}
#| label: setup-model-fitting
#| include: TRUE

# For data manipulation (loads dplyr, tidyr, purrr, tibble, etc.)
library(tidyverse)

# For file path management
library(here)

# The core package for Bayesian hierarchical modeling
library(brms)

# For model comparison and cross-validation (LOO)
library(loo)

# For creating interactive HTML tables
library(DT)


```

## Configuring Parallel Processing

```{r}

options(mc.cores = parallel::detectCores(), brms.backend = "cmdstanr")

```

::: {.callout-tip appearance="simple" icon="false"}
Threading can accelerate computation by distributing sampling across CPU cores. This requires the cmdstanr package and a working CmdStan installation.

```{r}
#| code-fold: FALSE
#| eval: false
#| label: Set up CmdStan and confirm proper installation
cmdstanr::install_cmdstan()
cmdstanr::check_cmdstan_toolchain()
cmdstanr::cmdstan_version()

```
:::

# Location–scale models of bloom phenology

These models tested whether both the mean timing and variability of bloom onset (days since the winter solstice) changed over time.

::: {.callout-note title="Variable definitions"}
-   delta5: Bloom onset date (days since winter solstice) at the 5% threshold.

-   delta15: Bloom onset date (days since winter solstice) at the 15% threshold.

-   delta25: Bloom onset date (days since winter solstice) at the 25% threshold.

-   TIME: Calendar year, centered (not scaled).
:::

## Data transformation

Bloom onset dates were converted to days since the previous winter solstice.

```{r}
#| label: bloom-data
#| eval: false
dat<-read.csv(here("..","data","bloom_dates.csv"))
dat <- dat %>%
  mutate(
    # Convert dates
    bloom_start_thr5_as_date = as.Date(bloom_start_thr5, format = "%d/%m/%Y"),
    bloom_start_thr15_as_date = as.Date(bloom_start_thr15, format = "%d/%m/%Y"),
    bloom_start_thr25_as_date = as.Date(bloom_start_thr25, format = "%d/%m/%Y"),
    winter_start_date = as.Date(paste0(season_year - 1, "-12-21")),

    # Calculate days since solstice
    delta5 = as.numeric(difftime(bloom_start_thr5_as_date, winter_start_date, units = "days")),
    delta15 = as.numeric(difftime(bloom_start_thr15_as_date, winter_start_date, units = "days")),
    delta25 = as.numeric(difftime(bloom_start_thr25_as_date, winter_start_date, units = "days")),

    # Factor year
    TIME = as.numeric(scale(season_year, center = TRUE, scale = FALSE)))%>%
  select( -starts_with("bloom_start_thr"), -winter_start_date)



```

## Model function

A wrapper function was used to fit Gaussian location–scale models for each bloom threshold (5%, 15%, 25%), with `TIME` included in both the location (μ) and scale (σ) submodels.

```{r}
#| label: bloom-function
#| eval: false
run_location_scale_models <- function(response_vars, data) {
  if (!dir.exists(here("Rdata"))) dir.create(here("Rdata"))
  
  purrr::walk(response_vars, function(var) {
    model_formula <- bf(
      as.formula(paste(var, "~ 1 + TIME")),
      sigma ~ 1 + TIME
    )
    
    file_path <- here("Rdata", paste0("model_", var, ".rds"))
    model_family <- gaussian()
    model_priors <- get_prior(model_formula, data = data, family = model_family)
    
    fit <- tryCatch({
      brm(
        formula = model_formula,
        data = data,
        prior = model_priors,
        family = model_family,
        iter = 6000, warmup = 2000, chains = 4, seed = 123, init = 0,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        file = file_path,
        file_refit = "on_change"
      )
    }, error = function(e) {
      cat("Error fitting model for:", var, "\n")
      cat("Error message:", e$message, "\n")
      return(NULL)
    })
    
    if (!is.null(fit)) cat("Saved model for:", var, "to", file_path, "\n")
  })
}
```

## Model execution

```{r}
#| label: bloom-ex
#| eval: false

# Bloom onset (delta variables)
delta_vars <- c("delta5", "delta15", "delta25")
run_location_scale_models(response_vars = delta_vars, data = dat)
```

# Hierarchical Bayesian models of reproductive performance

## Data preparation

The primary breeding dataset was loaded and pre-processed separately for females and males.

Key preprocessing steps included:

Type conversion of predictors to numeric

Encoding reproductive outcome (`FLEDS`) as an ordered factor for cumulative ordinal models

Centering (not scaling) continuous predictors (`TIME`, `X5MISMATCH`, `AGE`, `AFR`, `LONGEVITY`)

Converting `YEAR` and `RING` to factors for use as grouping variables

::: {.callout-note title="Variable definitions"}
-   `X5MISMATCH`: Phenological mismatch (days) using the 5% bloom threshold

-   `FLEDS`: Number of fledged chicks (ordinal response: 0–3)

-   `TIME`: Calendar year, centered (not scaled)

-   `ZX5MISMATCH`: Centered phenological mismatch (5% threshold)

-   `ABSMATCH`: Centered absolute mismatch (5% threshold)

-   `AGE`: Age, centered

-   `AFR`: Age at first reproduction, centered

-   `LONGEVITY`: Age at last observation (proxy for lifespan), centered

-   `YEAR`: Calendar year (factor)

-   `RING`: Individual identifier (factor)
:::

::: panel-tabset
## Load female data

```{r}
#| label: Data preparation females
#| eval: false
dat <- read.csv(here("..","data", "Lay_date_mismatches_250702.csv")) %>%
  subset(SEX == 1) %>%
  mutate(
    # First, ensure columns that will be used for scaling are numeric
    across(c(X5MISMATCH, AGE, AFR, LONGEVITY), as.numeric),
    
    # Second, create new scaled variables from existing columns
    TIME = scale(YEAR, center = TRUE, scale = FALSE)[,1],
    ZX5MISMATCH = scale(X5MISMATCH, center = TRUE, scale = FALSE)[,1],
    ABSMATCH = scale(abs(X5MISMATCH), center =TRUE, scale =FALSE)[,1],
    # Third, scale the other numeric predictors
    AGE = scale(AGE, center = TRUE, scale = FALSE)[,1],
    AFR = scale(AFR, center = TRUE, scale = FALSE)[,1],
    LONGEVITY = scale(LONGEVITY, center = TRUE, scale = FALSE)[,1],
    
    # Finally, convert grouping and outcome variables to factors
    YEAR = as.factor(YEAR),
    RING = as.factor(RING),
    FLEDS = factor(FLEDS, levels = 0:3, ordered = TRUE),
    
  )
```

## Load male data

```{r}
#| label: Data preparation male
#| eval: false
dat0 <- read.csv(here("..","data", "Lay_date_mismatches_250702.csv")) %>%
  subset(SEX ==0) %>%
  mutate(
    across(c(X5MISMATCH, AGE, AFR, LONGEVITY), as.numeric),
    
    TIME = scale(YEAR, center = TRUE, scale = FALSE)[,1],
    ZX5MISMATCH = scale(X5MISMATCH, center = TRUE, scale = FALSE)[,1],
    
    AGE = scale(AGE, center = TRUE, scale = FALSE)[,1],
    AFR = scale(AFR, center = TRUE, scale = FALSE)[,1],
    LONGEVITY = scale(LONGEVITY, center = TRUE, scale = FALSE)[,1],
    
    YEAR = as.factor(YEAR),
    RING = as.factor(RING),
    FLEDS = factor(FLEDS, levels = 0:3, ordered = TRUE),
    
  )
```
:::

## Models syntax

Two model families were used:

-   Gaussian models for continuous responses (e.g., phenological mismatch)

-   Cumulative ordinal probit models for ordered categorical responses (e.g., number of fledglings) 
::: panel-tabset 
## Gaussian models (Phenological mismatch)

```{r}
#| label: Gaussian models
#| echo: false
dfg <- read.csv(here("summaries","gaussian.csv"))
DT::datatable(dfg, options = list(pageLength = 5))
```

## Ordinal models (Number of fledglings)

```{r}
#| label: Ordinal models
#| echo: false
dfo <- read.csv(here("summaries","ordinal.csv"))
DT::datatable(dfo, options = list(pageLength = 5))
```
:::
# Automated model fitting workflow

To systematically evaluate candidate models, we used a custom function (`run_analysis_workflow()`) that automates model fitting, storage, and model comparison.

The workflow:

-   iterates over candidate fixed- and random-effects structures,

-   supports both Gaussian and ordinal families,

-   optionally models auxiliary parameters (sigma for Gaussian models; disc for ordinal models),

-   skips models that were already fitted,

-   saves fitted models and leave-one-out cross-validation (`LOO`) objects, and

-   uses `tryCatch()` so individual failures do not halt the full workflow.

::: panel-tabset
## Define Fixed Effects Structures

```{r}
#| label: Fixed Effects 
#| eval: false

FIXED_EFFECTS_ORDINAL_FULL  <- "ZX5MISMATCH + TIME + AGE + I(AGE^2) + AFR + LONGEVITY"

FIXED_EFFECTS_ORDINAL_QUADMATCH  <- "ZX5MISMATCH + I(ZX5MISMATCH^2) + TIME + AGE + I(AGE^2) + AFR + LONGEVITY"

FIXED_EFFECTS_ORDINAL_ABSMATCH  <- "ABSMATCH + TIME + AGE + I(AGE^2) + AFR + LONGEVITY"


FIXED_EFFECTS_GAUSSIAN      <- "AGE + I(AGE^2) + TIME + AFR + LONGEVITY"
```

## Function

```{r}
#| label: Custom function to fit hierarchical Bayesian models (females)
#| eval: false
# 2. MASTER WORKFLOW FUNCTION
# ==============================================================================
#' A flexible function to run various brms analysis workflows.
run_analysis_workflow <- function(analysis_name, response_var, fixed_effects, model_family, data, model_subset_key=NULL) {
  
  message(paste0("\n\n\n###--- Starting Analysis: '", analysis_name, "' ---###"))
  
  output_dir <- here("Rdata", analysis_name)
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
  
  is_gaussian <- model_family$family == "gaussian"
  
  # --- Define Model Specifications based on analysis type ---
  model_specs <- if (is_gaussian) {
    tribble(
      ~name,                      ~mean_random,               ~aux_formula,
      "m1",            "(1|RING) + (1|YEAR)",      NA, 
      "m2",       "(1+AGE|RING) + (1|YEAR)",  NA,
      "m3",      "(1+TIME|RING) + (1|YEAR)", NA,
      "m4",       "(1|q|RING) + (1|YEAR)",    ~ 1 + (1|q|RING),
      "m5",      "(1+AGE|q|RING) + (1|YEAR)",    ~ 1 + (1|q|RING),
      "m6",     "(1+TIME|q|RING) + (1|YEAR)",   ~ 1 + (1|q|RING),
      "m7",    "(1|q|RING) + (1|YEAR)",    ~ 1 + AGE + TIME + (1|q|RING),
      "m8",   "(1+AGE|q|RING) + (1|YEAR)",    ~ 1 + AGE + TIME + (1|q|RING),
      "m9",  "(1+TIME|q|RING) + (1|YEAR)",   ~ 1 + AGE + TIME + (1|q|RING),
      "m10",      "(1|q|RING) + (1|YEAR)",    ~ 1 + AGE + I(AGE^2) + TIME + (1|q|RING),
      "m11",     "(1+AGE|q|RING) + (1|YEAR)",    ~ 1 + AGE + I(AGE^2) + TIME + (1|q|RING),
      "m12",    "(1+TIME|q|RING) + (1|YEAR)",   ~ 1 + AGE + I(AGE^2) + TIME + (1|q|RING)
    )
  } else {
    tribble(
      ~name,                          ~mean_random,                          ~aux_formula,
    #LINEAR MISMATCH  
      "m1",                "(1|RING) + (1|YEAR)",                 ~ 1,
      "m2",           "(1+AGE|RING) + (1|YEAR)",             ~ 1,
      "m3",          "(1+TIME|RING) + (1|YEAR)",            ~ 1,
      "m4",      "(1+ZX5MISMATCH|RING) + (1|YEAR)",     ~ 1,
      "m5",            "(1|q|RING) + (1|YEAR)",               ~ 1 + (1|q|RING),
      "m6",           "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + (1|q|RING),
      "m7",          "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + (1|q|RING),
      "m8",      "(1+ZX5MISMATCH|q|RING) + (1|YEAR)",   ~ 1 + (1|q|RING),
      "m9",         "(1|q|RING) + (1|YEAR)",               ~ 1 + ZX5MISMATCH + AGE + TIME + (1|q|RING),
      "m10",        "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + ZX5MISMATCH + AGE + TIME + (1|q|RING),
      "m11",       "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + ZX5MISMATCH + AGE + TIME + (1|q|RING),
      "m12",   "(1+ZX5MISMATCH|q|RING) + (1|YEAR)",   ~ 1 + ZX5MISMATCH + AGE + TIME + (1|q|RING),
      "m13",           "(1|q|RING) + (1|YEAR)",               ~ 1 + ZX5MISMATCH + AGE + I(AGE^2) + TIME + (1|q|RING),
      "m14",          "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + ZX5MISMATCH + AGE + I(AGE^2) + TIME + (1|q|RING),
      "m15",         "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + ZX5MISMATCH + AGE + I(AGE^2) + TIME + (1|q|RING),
      "m16",     "(1+ZX5MISMATCH|q|RING) + (1|YEAR)",   ~ 1 + ZX5MISMATCH + AGE + I(AGE^2) + TIME + (1|q|RING),
      
      
# QUAD MISMATCH      
      "Q1",                "(1|RING) + (1|YEAR)",                 ~ 1,
      "Q2",           "(1+AGE|RING) + (1|YEAR)",             ~ 1,
      "Q3",          "(1+TIME|RING) + (1|YEAR)",            ~ 1,
      "Q4",      "(1+ZX5MISMATCH|RING) + (1|YEAR)",     ~ 1,
      "Q5",            "(1|q|RING) + (1|YEAR)",               ~ 1 + (1|q|RING),
      "Q6",           "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + (1|q|RING),
      "Q7",          "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + (1|q|RING),
      "Q8",      "(1+ZX5MISMATCH|q|RING) + (1|YEAR)",   ~ 1 + (1|q|RING),
       "Q9",         "(1|q|RING) + (1|YEAR)",               ~ 1 + ZX5MISMATCH + I(ZX5MISMATCH^2) + AGE + TIME + (1|q|RING),
      "Q10",        "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + ZX5MISMATCH + I(ZX5MISMATCH^2) +AGE + TIME + (1|q|RING),
      "Q11",       "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + ZX5MISMATCH + I(ZX5MISMATCH^2) +AGE + TIME + (1|q|RING),
      "Q12",   "(1+ZX5MISMATCH|q|RING) + (1|YEAR)",   ~ 1 + ZX5MISMATCH + I(ZX5MISMATCH^2) +AGE + TIME + (1|q|RING),
      "Q13",           "(1|q|RING) + (1|YEAR)",               ~ 1 + ZX5MISMATCH + I(ZX5MISMATCH^2) +AGE + I(AGE^2) + TIME + (1|q|RING),
      "Q14",          "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + ZX5MISMATCH + I(ZX5MISMATCH^2) +AGE + I(AGE^2) + TIME + (1|q|RING),
      "Q15",         "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + ZX5MISMATCH + I(ZX5MISMATCH^2) +AGE + I(AGE^2) + TIME + (1|q|RING),
      "Q16",     "(1+ZX5MISMATCH|q|RING) + (1|YEAR)",   ~ 1 + ZX5MISMATCH + I(ZX5MISMATCH^2) +AGE + I(AGE^2) + TIME + (1|q|RING),


       "Q17",         "(1|q|RING) + (1|YEAR)",               ~ 1 + ZX5MISMATCH +AGE + TIME + (1|q|RING),
      "Q18",        "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + ZX5MISMATCH +AGE + TIME + (1|q|RING),
      "Q19",       "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + ZX5MISMATCH  +AGE + TIME + (1|q|RING),
      "Q20",   "(1+ZX5MISMATCH|q|RING) + (1|YEAR)",   ~ 1 + ZX5MISMATCH  +AGE + TIME + (1|q|RING),
      "Q21",           "(1|q|RING) + (1|YEAR)",               ~ 1 + ZX5MISMATCH  +AGE + I(AGE^2) + TIME + (1|q|RING),
      "Q22",          "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + ZX5MISMATCH  +AGE + I(AGE^2) + TIME + (1|q|RING),
      "Q23",         "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + ZX5MISMATCH  +AGE + I(AGE^2) + TIME + (1|q|RING),
      "Q24",     "(1+ZX5MISMATCH|q|RING) + (1|YEAR)",   ~ 1 + ZX5MISMATCH  +AGE + I(AGE^2) + TIME + (1|q|RING),
      
      
      
      
#ABS MISMATCH      
      "ABS1",                "(1|RING) + (1|YEAR)",                 ~ 1,
      "ABS2",           "(1+AGE|RING) + (1|YEAR)",             ~ 1,
      "ABS3",          "(1+TIME|RING) + (1|YEAR)",            ~ 1,
      "ABS4",      "(1+ABSMATCH|RING) + (1|YEAR)",     ~ 1,
      "ABS5",            "(1|q|RING) + (1|YEAR)",               ~ 1 + (1|q|RING),
      "ABS6",           "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + (1|q|RING),
      "ABS7",          "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + (1|q|RING),
      "ABS8",      "(1+ABSMATCH|q|RING) + (1|YEAR)",   ~ 1 + (1|q|RING),
       "ABS9",         "(1|q|RING) + (1|YEAR)",               ~ 1 + ABSMATCH + AGE + TIME + (1|q|RING),
      "ABS10",        "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + ABSMATCH+ AGE + TIME + (1|q|RING),
      "ABS11",       "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + ABSMATCH + AGE + TIME + (1|q|RING),
      "ABS12",   "(1+ABSMATCH|q|RING) + (1|YEAR)",   ~ 1 + ABSMATCH+ AGE + TIME + (1|q|RING),
      "ABS13",           "(1|q|RING) + (1|YEAR)",               ~ 1 + ABSMATCH + AGE + I(AGE^2) + TIME + (1|q|RING),
      "ABS14",          "(1+AGE|q|RING) + (1|YEAR)",           ~ 1 + ABSMATCH + AGE + I(AGE^2) + TIME + (1|q|RING),
      "ABS15",         "(1+TIME|q|RING) + (1|YEAR)",          ~ 1 + ABSMATCH + AGE + I(AGE^2) + TIME + (1|q|RING),
      "ABS16",     "(1+ABSMATCH|q|RING) + (1|YEAR)",   ~ 1 + ABSMATCH+ AGE + I(AGE^2) + TIME + (1|q|RING),
      
      
      
      
      
      
      
      
      
      
      
      
    )
  }
  
if (!is.null(model_subset_key)) {
  message(paste0("Filtering models for subset: '", model_subset_key, "'"))
  
  if (model_subset_key == "linear") {
   # Keep only models that DON'T start with "Q_" or "ABS_"
   model_specs <- model_specs %>%
    filter(!grepl("^Q", name) & !grepl("^ABS", name))
  } else if (model_subset_key == "quad") {
   # Keep only models that DO start with "Q_"
   model_specs <- model_specs %>%
    filter(grepl("^Q", name))
  } else if (model_subset_key == "abs") {
   # Keep only models that DO start with "ABS_"
   model_specs <- model_specs %>%
    filter(grepl("^ABS", name))
  }
  
  # Safety check
  if (nrow(model_specs) == 0) {
   warning(paste0("No models found for subset key: '", model_subset_key, "'. Check your model names and key."))
   return(invisible(NULL)) # Stop if no models to run
  }
 }
  
  
  # --- Internal function to fit a single model ---
  fit_and_save_model <- function(name, mean_random, aux_formula, ...) {
    model_path <- file.path(output_dir, paste0("female_fit_", name, ".rds"))
    loo_path <- file.path(output_dir, paste0("female_loo_", name, ".rds"))
    
    if (file.exists(model_path)) {
      message(paste0("Skipping '", name, "': Model file already exists."))
      return(invisible(NULL))
    }
    
    message(paste0("\n--- Fitting model: ", name, " ---"))
    
    formula_str <- paste(response_var, "~", fixed_effects, "+", mean_random)
    
    # --- CORRECTED LOGIC ---
    # Use inherits() for a robust check if aux_formula is a real formula or NA.
    brm_formula <- if (!inherits(aux_formula, "formula")) {
      bf(formula_str, nl = FALSE)
    } else if (is_gaussian) {
      bf(formula_str, sigma = aux_formula, nl = FALSE)
    } else {
      bf(formula_str, disc = aux_formula, nl = FALSE)
    }
    
    model_priors <- get_prior(brm_formula, data = data, family = model_family)
    
    fit <- tryCatch({
      brm(
        formula = brm_formula, data = data, prior = model_priors, family = model_family,
        iter = 6000, warmup = 2000, chains = 4, seed = 123, init = 0,backend = "cmdstanr",threads = threading(12),
        control = list(adapt_delta = 0.99, max_treedepth = 15)
      )
    }, error = function(e) { message(paste0("ERROR fitting '", name, "': ", e$message)); NULL })
    
    if (!is.null(fit)) {
      saveRDS(fit, model_path)
      loo_result <- tryCatch(loo(fit), error = function(e) { message(paste0("ERROR with LOO: ", e$message)); NULL })
      if (!is.null(loo_result)) saveRDS(loo_result, loo_path)
    }
  }
  
  # --- Run all models and compare LOO results ---
  pwalk(model_specs, fit_and_save_model)
  

}
```
:::

# Executing the workflow for all analyses

The workflow was run separately for males (dat0) and females (dat) for:

-   Phenological mismatch (Gaussian response: `X5MISMATCH`)

-   Fledging success (ordinal response: `FLEDS`)

-   Each candidate model set was fitted and saved for subsequent model comparison.

::::: panel-tabset
## Analyses for males

The workflow was first run using the male dataset (dat0).

::: panel-tabset
## Phenological mismatch Analysis

```{r}
#| label: Execute pheno males
#| eval: false
run_analysis_workflow(
  analysis_name = "pheno_mismatch",
  response_var = "X5MISMATCH",
  fixed_effects = FIXED_EFFECTS_GAUSSIAN,
  model_family = gaussian(),
  data = dat0
)
```

## Number of fledglings Analysis

Here, the response variable (\``FLEDS`) is an ordered categorical outcome (0, 1, 2, or 3 fledglings), so models were fitted using a cumulative probit family

```{r}
#| label: Execute fleds males
#| eval: false
run_analysis_workflow(
  analysis_name = "fleds_full",
  response_var = "FLEDS",
  fixed_effects = FIXED_EFFECTS_ORDINAL_FULL,
  model_family = cumulative(link = "probit"),
  data = dat0,
  model_subset_key = "linear"
)

run_analysis_workflow(
 analysis_name = "fleds_full",
 response_var = "FLEDS",
 fixed_effects = FIXED_EFFECTS_ORDINAL_QUADMATCH,
 model_family = cumulative(link = "probit"),
 data = dat0,
 model_subset_key = "quad" 
)

run_analysis_workflow(
 analysis_name = "fleds_full",
 response_var = "FLEDS",
 fixed_effects = FIXED_EFFECTS_ORDINAL_ABSMATCH ,
 model_family = cumulative(link = "probit"),
 data = dat0,
 model_subset_key = "abs" 
)
```
:::

## Analyses for females

The same workflow was then run using the female dataset (dat)

::: panel-tabset
## Phenological mismatch Analysis

```{r}
#| label: Execute pheno females
#| eval: false
run_analysis_workflow(
  analysis_name = "pheno_mismatch",
  response_var = "X5MISMATCH",
  fixed_effects = FIXED_EFFECTS_GAUSSIAN,
  model_family = gaussian(),
  data = dat
)
```

## Number of fledglings Analysis

```{r}
#| label: Execute fleds females
#| eval: false
run_analysis_workflow(
  analysis_name = "fleds_full",
  response_var = "FLEDS",
  fixed_effects = FIXED_EFFECTS_ORDINAL_FULL,
  model_family = cumulative(link = "probit"),
  data = dat,
  model_subset_key = "linear"
)

run_analysis_workflow(
 analysis_name = "fleds_full",
 response_var = "FLEDS",
 fixed_effects = FIXED_EFFECTS_ORDINAL_QUADMATCH,
 model_family = cumulative(link = "probit"),
 data = dat,
 model_subset_key = "quad" 
)

run_analysis_workflow(
 analysis_name = "fleds_full",
 response_var = "FLEDS",
 fixed_effects = FIXED_EFFECTS_ORDINAL_ABSMATCH ,
 model_family = cumulative(link = "probit"),
 data = dat,
 model_subset_key = "abs" 
)

```
:::
:::::

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
