---
title: "S4. Model diagnostics and selection"
---

```{r}
#| label: setup-model-diagnostics
#| include: false

# For data manipulation (loads dplyr, tibble, etc.)
library(tidyverse)

# For file path management
library(here)

# For accessing model diagnostics (e.g., divergences)
library(brms)

# For summarizing posterior draws (e.g., for ESS checks)
library(posterior)

# For model comparison using LOO-CV
library(loo)

# For creating interactive HTML tables
library(DT)
```

# Model diagnostics

This section documents the R code used to evaluate model diagnostics and perform model comparisons for the Bayesian analyses reported in the main text. The workflow has two components:

Model diagnostics: automated checks for common fitting issues (e.g., divergent transitions and low effective sample size, ESS)

Model comparison: evaluation of predictive performance using leave-one-out cross-validation (LOO-CV)

These procedures were applied to the pheno_mismatch and fleds_full analysis sets
::: {.callout-important appearance="simple"}
Diagnostics for bloom onset location–scale models were inspected directly from model summaries because only one model was fitted per threshold.
:::

## Diagnostic and comparison functions

::: panel-tabset
## Function 1: `create_diagnostic_summary()`

This function automates convergence and sampling-efficiency checks for Bayesian models. It identifies saved model files, loads them, and performs one of two checks:

- Divergences: counts divergent transitions, which may indicate biased posterior exploration.

- ESS: evaluates bulk and tail effective sample sizes, flagging parameters with insufficient independent samples (< 400). We used a conservative threshold of 400 for both bulk and tail ESS to flag potentially unstable estimates for further inspection

Results are returned as a summary table for each model (separated by sex), including the diagnostic outcome and any flagged parameters.

```{r}
#| label: Function 1
#| eval: false
#| code-fold: true
create_diagnostic_summary <- function(analysis_name, check_type = "divergences", ess_threshold = 400) {
  # --- Load required libraries ---
  if (!require(here)) {install.packages("here"); library(here)}
  if (!require(brms)) {install.packages("brms"); library(brms)}
  if (!require(dplyr)) {install.packages("dplyr"); library(dplyr)}

  # --- 1. Find all male and female model files ---
  output_dir <- here("Rdata", analysis_name)
  model_files <- list.files(output_dir, pattern = "^(male|female)_fit_.*\\.rds$", full.names = TRUE)

  if (length(model_files) == 0) {
    message("No model files found in ", output_dir)
    return(invisible(NULL))
  }

  # --- 2. Initialize a list to store results ---
  results_list <- list()

  # --- 3. Loop through files, run checks, and store results ---
  for (model_path in model_files) {
    fit <- readRDS(model_path)
    file_name <- basename(model_path)
    
    # Extract sex and model name
    sex <- sub("_fit_.*", "", file_name)
    model_name <- sub("^(male|female)_fit_(.*)\\.rds$", "\\2", file_name)

    message(paste("Checking:", sex, model_name))

    # --- Run the requested check ---
    if (tolower(check_type) == "divergences") {
      nuts_params <- nuts_params(fit)
      check_value <- sum(subset(nuts_params, Parameter == "divergent__")$Value)
      problem_params <- ifelse(check_value > 0, "Divergences present", NA_character_)

    } else if (tolower(check_type) == "ess") {
      if (!require(posterior)) {install.packages("posterior"); library(posterior)}
      draws_summary <- posterior::summarise_draws(fit)
      low_ess_params <- filter(draws_summary, ess_bulk < ess_threshold | ess_tail < ess_threshold)
      check_value <- nrow(low_ess_params)
      problem_params <- ifelse(check_value > 0, paste(low_ess_params$variable, collapse = ", "), NA_character_)

    } else {
      stop("Invalid 'check_type'. Please use 'divergences' or 'ess'.")
    }

    # Store the results for the current model
    results_list[[length(results_list) + 1]] <- tibble(
      Sex = sex,
      Model = model_name,
      Check_Result = check_value,
      Problematic_Parameters = problem_params
    )
  }

  # --- 4. Combine list into a final table and return ---
  final_table <- bind_rows(results_list)

  # Rename the Check_Result column to be more descriptive
  if (tolower(check_type) == "divergences") {
    final_table <- rename(final_table, Divergences = Check_Result)
  } else {
    final_table <- rename(final_table, Low_ESS_Count = Check_Result)
  }
  
  return(final_table)
}
```

## Function 2: `list_analysis_models()`

This function inventories pre-computed LOO objects for a given analysis and returns a data frame listing the available models (sex, model name, file path) for comparison.
```{r}
#| label: Function 2
#| eval: false
list_analysis_models <- function(analysis_name) {
  if (!require(here)) {install.packages("here"); library(here)}
  if (!require(dplyr)) {install.packages("dplyr"); library(dplyr)}
  
  output_dir <- here("Rdata", analysis_name)
  all_loo_files <- list.files(output_dir, pattern = "^(male|female)_loo_.*\\.rds$", full.names = TRUE)
  
  if (length(all_loo_files) == 0) {
    message("No 'male' or 'female' LOO files found in ", output_dir)
    return(tibble())
  }
  
  model_inventory <- tibble(file_path = all_loo_files) %>%
    mutate(
      file_name = basename(file_path),
      sex = sub("_loo_.*", "", file_name),
      model_name = sub("^(male|female)_loo_(.*)\\.rds$", "\\2", file_name)
    ) %>%
    select(sex, model_name, file_path)
    
  return(model_inventory)
}
```

## Function 3: `compare_models_by_sex()`
This function compares models using the loo package. Models are grouped by sex, optionally excluding problematic fits, and ranked by expected log predictive density (ELPD). The comparison output includes differences in ELPD, standard errors, and LOO Information Criterion (LOOIC).

```{r}
#| label: Function 3
#| eval: false
compare_models_by_sex <- function(model_inventory, exclude_models = NULL) {
  if (!require(loo)) {install.packages("loo"); library(loo)}
  if (!require(dplyr)) {install.packages("dplyr"); library(dplyr)}

  comparison_results <- list()

  for (current_sex in c("male", "female")) {
    cat(paste0("\n\n###--- LOO Comparison for ", toupper(current_sex), " Models ---###\n\n"))

    # Filter the inventory for the current sex
    sex_models_df <- model_inventory %>% filter(sex == current_sex)

    # Get the specific list of models to exclude
    models_to_exclude <- exclude_models[[current_sex]]
    
    if (!is.null(models_to_exclude) && nrow(sex_models_df) > 0) {
      sex_models_df <- sex_models_df %>% filter(!model_name %in% models_to_exclude)
      message(paste("For", current_sex, ", excluding models:", paste(models_to_exclude, collapse=", ")))
    }

    if (nrow(sex_models_df) < 2) {
      message(paste("Not enough models to compare for '", current_sex, "' (need at least 2). Skipping."))
      next
    }

    # Load the relevant LOO objects into a named list
    loo_list <- setNames(
      lapply(sex_models_df$file_path, readRDS), 
      sex_models_df$model_name
    )

    message(paste0("Comparing ", length(loo_list), " '", current_sex, "' models..."))
    
    # Perform the comparison and print it
    comparison <- loo_compare(x = loo_list)
    print(comparison)
    
    comparison_results[[current_sex]] <- comparison
  }

  invisible(comparison_results)
}
```
:::

# Application to Analyses



The following sections apply these functions to each analysis set.

::::: panel-tabset
## Diagnostic Checks

Both divergent transitions and low ESS values were checked across all models in this analysis set.

::: panel-tabset
## Divergent transitions

```{r}
#| label: Check for divergences PM
#| eval: false
# Check for divergent transitions
divergence_summaryPM <- create_diagnostic_summary("pheno_mismatch", check_type = "divergences")
print(divergence_summaryPM)
```

```{r}
#| label: DIVERGENCIES PM
#| echo: false
divergence_summaryPM<-read.csv(here("summaries","divergence_summaryPM.csv"),header=T)
DT::datatable(divergence_summaryPM, options = list(pageLength = 5))

```

## Effective Sample Size (ESS)

```{r}
#| label: Check for divergences and low ESS pheno_mismatch
#| eval: false
# Check for low Effective Sample Size (ESS)
ess_summaryPM <- create_diagnostic_summary("pheno_mismatch", check_type = "ess", ess_threshold = 400)
print(ess_summaryPM)
```

```{r}
#| label: ESS PM
#| echo: false
ess_summaryPM<-read.csv(here("summaries","ess_summaryPM.csv"),header=T)
DT::datatable(ess_summaryPM, options = list(pageLength = 5))
```
:::

## Model comparison
Pre-computed LOO objects were compared separately for males and females after excluding models flagged as problematic in diagnostic checks.
```{r}
#| label: Model comparison pheno_mismatch
#| eval: false
# Create an inventory of all LOO objects for the analysis
model_inventoryPM <- list_analysis_models("pheno_mismatch")

# Define models to exclude from comparison (e.g., due to divergences or low ESS)
bad_models_list <- list(
  male = c("m12","m6","m9"),
   female = c("m12","m6","m9")
)
# Run the LOO comparison, excluding the problematic models
all_comparisonsPM <- compare_models_by_sex(model_inventoryPM, exclude_models = bad_models_list)
```

::: panel-tabset
## Female summary

```{r}
#| label: MODELS PM females
#| echo: false
all_comparisonsPMf<-read.csv(here("summaries","all_comparisonsPMf.csv"),header=T) %>%
  select(Model, elpd_diff, se_diff)
DT::datatable(all_comparisonsPMf, options = list(pageLength = 5))

```

## Male summary

```{r}
#| label: MODELS PM males
#| echo: false
all_comparisonsPMm<-read.csv(here("summaries","all_comparisonsPMm.csv"),header=T) %>%
  select(Model, elpd_diff, se_diff)
DT::datatable(all_comparisonsPMm, options = list(pageLength = 5))

```
:::
:::::

## Number of fledglings (`fleds_full`)

::::: panel-tabset
## Diagnostic Checks

::: panel-tabset
## Divergent transitions

```{r}
#| label: Check for divergences  fleds_full
#| eval: false

divergence_summaryFS <- create_diagnostic_summary("fleds_full", check_type = "divergences")
DT::datatable(divergence_summaryFS, options = list(pageLength = 5))


```

```{r}
#| label: divergences fleds_full
#| echo: false
divergence_summaryFS <-read.csv(here("summaries","divergence_summaryFS.csv"),header=T)

DT::datatable(divergence_summaryFS, options = list(pageLength = 5))
```

## Low ESS

```{r}
#| label: Check for low ESS fleds_full
#| eval: false
ess_summaryFS <- create_diagnostic_summary("fleds_full", check_type = "ess", ess_threshold = 400)
DT::datatable(ess_summaryFS, options = list(pageLength = 5))
```

```{r}
#| label: ESS fleds_full
#| echo: false
ess_summaryFS<-read.csv(here("summaries","ess_summaryFS .csv"),header=T)
DT::datatable(ess_summaryFS, options = list(pageLength = 5))
```
:::

## Model comparison

```{r}
#| label: Model comparison fleds_full
#| eval: false

# Create an inventory of all LOO objects for the analysis
model_inventoryFS <- list_analysis_models("fleds_full")
print(model_inventoryFS)

# Define models to exclude from comparison
bad_models_listFS <- list(
  male = c("ABS10","ABS16","Q3"
           ),
   female = c("ABS3","m3","Q9","Q3","Q6"
              )
)

# Run the LOO comparison
all_comparisonsFS <- compare_models_by_sex(model_inventoryFS, exclude_models = bad_models_listFS)
```

::: panel-tabset
## Female summary

```{r}
#| label: MODELS FS female
#| echo: false
all_comparisonsFSf <-read.csv(here("summaries","all_comparisonsFSf.csv"),header=T) %>%
  select(Model, elpd_diff, se_diff)
DT::datatable(all_comparisonsFSf, options = list(pageLength = 5))
```

## Male summary

```{r}
#| label: MODELS FS male
#| echo: false
all_comparisonsFSm <-read.csv(here("summaries","all_comparisonsFSm.csv"),header=T) %>%
  select(Model, elpd_diff, se_diff)
DT::datatable(all_comparisonsFSm, options = list(pageLength = 5))
```
:::
:::::

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
