---
title: "S5. Model comparison"
---

```{r}
#| label: setup-model-comparison
#| include: false

# Load libraries needed for this chapter
pacman::p_load(here, dplyr, DT,loo)
```

This section reports model comparison results used to identify the best-supported model for each analysis.

All candidate models were evaluated using leave-one-out cross-validation (LOO-CV), which estimates out-of-sample predictive performance. Models are ranked by the difference in expected log predictive density (elpd_diff) relative to the top-ranked model.

-   The best-supported model in each table has `elpd_diff = 0`.

-   `se_diff` gives the standard error of the difference in ELPD, indicating uncertainty in the ranking \# Models syntax

::: panel-tabset
## Gaussian models (Phenological mismatch)

```{r}
#| label: Gaussian models
#| echo: false
dfg <- read.csv(here("summaries","gaussian.csv"))
DT::datatable(dfg, options = list(pageLength = 5))
```

## Ordinal models (Number of fledglings)

```{r}
#| label: Ordinal models
#| echo: false
dfo <- read.csv(here("summaries","ordinal.csv"))
DT::datatable(dfo, options = list(pageLength = 5))
```
:::

# Phenological Mismatch

The following tables summarize LOO-CV comparisons for Gaussian models of phenological mismatch, shown separately for females and males. 
::: panel-tabset 
## Females

```{r}
#| label: loo-comparison-pm-females
#| echo: false
#| tbl-cap: "LOO-CV comparison for female phenological mismatch models."

all_comparisonsPMf <- read.csv(here::here("summaries", "all_comparisonsPMf.csv"), header = TRUE) %>%
  select(Model, elpd_diff, se_diff)

DT::datatable(all_comparisonsPMf, options = list(pageLength = 10))
```

## Males

```{r}
#| label: loo-comparison-pm-males
#| echo: false
#| tbl-cap: "LOO-CV comparison for male phenological mismatch models."

all_comparisonsPMm <- read.csv(here::here("summaries", "all_comparisonsPMm.csv"), header = TRUE) %>%
  select(Model, elpd_diff, se_diff)

DT::datatable(all_comparisonsPMm, options = list(pageLength = 10))
```
:::
# Number of Fledglings

The following tables summarize LOO-CV comparisons for ordinal models of fledgling number, shown separately for females and males. 
::: panel-tabset 
## Females

```{r}
#| label: loo-comparison-fs-females
#| echo: false
#| tbl-cap: "LOO-CV comparison for female fledgling success models."

all_comparisonsFSf <- read.csv(here::here("summaries", "all_comparisonsFSf.csv"), header = TRUE) %>%
  select(Model, elpd_diff, se_diff)

DT::datatable(all_comparisonsFSf, options = list(pageLength = 10))
```

## Males

```{r}
#| label: loo-comparison-fs-males
#| echo: false
#| tbl-cap: "LOO-CV comparison for male fledgling success models."

all_comparisonsFSm <- read.csv(here::here("summaries", "all_comparisonsFSm.csv"), header = TRUE) %>%
  select(Model, elpd_diff, se_diff)

DT::datatable(all_comparisonsFSm, options = list(pageLength = 10))
```
:::
::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
