---
title: "S6. Full model summaries"
unnumbered: true
---

```{r}
#| label: setup-final-summaries
#| include: false

# Load libraries needed for this chapter
pacman::p_load(here, brms)
```

::: {.callout-note title="Variable definition"}
- `X5MISMATCH`: Phenological mismatch calculated using the 5% bloom threshold (days).

- `CLUTCH_SIZE`: Number of eggs laid in a nest (ordinal).

- `FLEDS`: Number of chicks successfully fledged (ordinal).

- `TIME`: Chronological time (calendar year), centered by subtracting the mean year.

- `ZX5MISMATCH`: Centered phenological mismatch (5% threshold).

- `AGE`: Age of the focal individual (years), centered.

- `AFR`: Age at first reproduction (years), centered.

- `LONGEVITY`: Age at last observation (proxy for lifespan; years), centered.

- `YEAR`: Calendar year of the observation (factor).

- `RING`: Individual identifier (factor).
:::

This section reports posterior summaries for the best-supported models from the primary analyses. Best-supported models for phenological mismatch and number of fledglings were selected using leave-one-out cross-validation (LOO-CV; see S5).


## Bloom timing (location–scale model)

```{r}
#| label: summary-bloom
#| eval: true
#| size: "footnotesize"


# Load the model
model_bloom <- readRDS(here::here("..","Rdata", "model_delta5.rds"))

# Print the full model summary
print(summary(model_bloom),digits = 4)

```


## Best-supported individual-based models
::::: panel-tabset
## Females

::: panel-tabset
## Phenological Mismatch

```{r}
#| label: summary-pm-female-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_pm_female_best <- readRDS(here::here("..","Rdata", "female_fit_m11.rds"))

# Print the full model summary
print(summary(model_pm_female_best),digits=4)

```

## Number of Fledglings

```{r}
#| label: summary-fs-female-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_fs_female_best <- readRDS(here::here("..","Rdata", "female_fit_Q17.rds"))

# Print the full model summary
print(summary(model_fs_female_best), digits =4)
```
:::

## Males

::: panel-tabset
## Phenological Mismatch

```{r}
#| label: summary-pm-male-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_pm_male_best <- readRDS(here::here("..","Rdata", "male_fit_m11.rds"))

# Print the full model summary
print(summary(model_pm_male_best),digits=4)
```

## Number of Fledglings

```{r}
#| label: summary-fs-male-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_fs_male_best <- readRDS(here::here("..","Rdata", "male_fit_Q17.rds"))

# Print the full model summary
print(summary(model_fs_male_best),digits=4)
```
:::
:::::

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
