---
title: "Code for figures"
subtitle: "Reproducible workflow for generating the main visualizations"
unnumbered: true
toc: true
toc-depth: 2
---

::: lead
This appendix provides the code used to generate the main figures in the manuscript. It includes setup, data preparation, model-based prediction workflows, plot construction, and export code for the final multi-panel figures.
:::

::: {.callout-note appearance="simple"}
Most chunks are included as a reproducible workflow record and are not executed during rendering of the website.
:::

```{r}
#| label: setup-pheno_results
#| include: false

pacman::p_load(MASS, 
               dplyr,
               ggplot2,broom.mixed,
               here,ggdist,patchwork,
               arm,
               tidyverse,
               tidybayes,
               brms,
               rstan,
               emmeans,
               ggthemes,
               bayesplot ,cmdstanr, extrafont, ggtext)
```

## Plotting utilities

This section defines reusable plotting objects to ensure a consistent visual style across figures.

### Custom theme

```{r}
#| label: custom-theme
#| eval: false
#| code-fold: true
#| code-summary: "Show custom plotting theme"
#| 
custom_theme <- function(show_legend = FALSE) {
  # Set legend position string
  legend_pos <- if (show_legend) "bottom" else "none"
  
  # Use theme_classic() as the base
  theme_classic(
    base_size = 10,
    base_family = "Arial"
  ) +
  theme(
    # --- Add minor ticks back in ---
    # This is the line that fixes your problem
    axis.ticks.minor = element_line(color = "black", linewidth = 0.25), 
    
    # --- Legend and margins ---
    legend.position = legend_pos,
    plot.margin = margin(5, 5, 5, 5),
    
    # --- Backgrounds (optional, theme_classic is not transparent) ---
    panel.background = element_rect(fill = 'transparent', colour = NA),
    plot.background = element_rect(fill = 'transparent', colour = NA),
    
    # --- Text and Titles (your original customizations) ---
    axis.title.x = element_text(margin = margin(t = 5)),
    axis.title.y = element_text(margin = margin(r = 5)),
    axis.text.x = element_text(margin = margin(t = 2)),
    axis.text.y = element_text(hjust = 1, margin = margin(r = 2)),
    plot.title = element_text(hjust = 0.5, face = "bold", size = rel(1.2), margin = margin(b = 8)),
    plot.subtitle = element_text(hjust = 0.5, size = rel(1.0), margin = margin(b = 8)),
    
    # --- Axis Lines (already in theme_classic, but good to keep) ---
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black")
  )
}
```

### Shared x-axis scale

```{r}
#| label: shared-x-axis-scale
#| eval: false
#| code-fold: true
#| code-summary: "Show shared x-axis scale"

# Create a reusable x-axis scale for all plots
my_x_scale <- scale_x_continuous(
  name = "Time (years)", # Set the axis title here
  breaks = seq(2003, 2023, 2),
  minor_breaks = seq(2003, 2023, 1),
  labels = function(breaks) {
    # This function shows the label only if the year is the first one (2003)
    # or is a multiple of 4 years from the start.
    ifelse((breaks - 2003) %% 4 == 0, as.character(breaks), "")
  },
  guide = guide_axis(minor.ticks = TRUE)
)
```

## Load data and fitted models

Data and pre-trained models were loaded from project files prior to figure generation.

::: panel-tabset
## Bloom onset

```{r}
#| label: load-bloom-data-model
#| eval: false
#| code-fold: true
#| code-summary: "Show bloom data and model loading"

bloom_dat_raw <- read.csv(here("..","data", "bloom_dates.csv"))
bloom_model <- readRDS(here("..","Rdata", "model_delta5.rds"))

```

## Females

```{r}
#| label: load-female-data-models
#| eval: false
#| code-fold: true
#| code-summary: "Show female data and model loading"

female_dat_raw <- read.csv(here("..","data", "Lay_date_mismatches_250702.csv"))
mismatch_modelf <- readRDS(here("..","Rdata", "female_fit_m11.rds"))
fledging_modelf <- readRDS(here("..","Rdata", "female_fit_Q17.rds"))
```

## Males

```{r}
#| label: load-male-data-models
#| eval: false
#| code-fold: true
#| code-summary: "Show male data and model loading"


male_dat_raw <- read.csv(here("..","data", "Lay_date_mismatches_250702.csv"))
mismatch_modelm <- readRDS(here("..","Rdata", "male_fit_m11.rds"))
fledging_modelm <- readRDS(here("..","Rdata", "male_fit_Q17.rds"))
```
:::

### Precomputed posterior draws

Posterior draws used for plotting are precomputed on `totoro` and saved as `.rds` files. This keeps local figure editing fast because plot chunks can load the draw objects directly instead of recalculating them from the fitted models.

```{r}
#| label: load-precomputed-draw-helper
#| eval: false
#| code-fold: true
#| code-summary: "Show precomputed draw loading helper"

precomputed_draw_dir <- here("..", "Rdata", "precomputed_draws")

read_precomputed_draw <- function(file) {
  readRDS(file.path(precomputed_draw_dir, file))
}
```

::: {.callout-note appearance="simple"}
The following folded chunks document the posterior-draw recipes that were run on `totoro` using `Rscripts/precompute_figure_draws.R`. They assume the preprocessing objects defined below are available and are kept with `eval: false` so the website records how each `.rds` object was generated, while the plotting chunks load the precomputed files from `Rdata/precomputed_draws`.
:::

```{r}
#| label: totoro-figure2-draw-recipes
#| eval: false
#| code-fold: true
#| code-summary: "Show totoro recipes for Figure 2 posterior draws"

mean_bloom_year <- mean(bloom_dat$season_year)
bloom_time_draws <- bloom_model %>%
  epred_draws(
    newdata = expand.grid(TIME = seq(min(bloom_dat$TIME), max(bloom_dat$TIME), length.out = 100)),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_year = TIME + mean_bloom_year)

mean_mismatch_yearf <- mean(dat_female_mismatch$ori_time)
mismatch_time_drawsf <- mismatch_modelf %>%
  epred_draws(
    newdata = expand.grid(
      TIME = seq(min(dat_female_mismatch$TIME), max(dat_female_mismatch$TIME), length.out = 100),
      AGE = mean(dat_female_mismatch$AGE),
      AFR = mean(dat_female_mismatch$AFR),
      LONGEVITY = mean(dat_female_mismatch$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_time = TIME + mean_mismatch_yearf)

mean_mismatch_yearm <- mean(dat_male_mismatch$ori_time)
mismatch_time_drawsm <- mismatch_modelm %>%
  epred_draws(
    newdata = expand.grid(
      TIME = seq(min(dat_male_mismatch$TIME), max(dat_male_mismatch$TIME), length.out = 100),
      AGE = mean(dat_male_mismatch$AGE),
      AFR = mean(dat_male_mismatch$AFR),
      LONGEVITY = mean(dat_male_mismatch$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_time = TIME + mean_mismatch_yearm)

mean_fledging_yearf <- mean(dat_female_fledging$ori_time)
fledging_time_drawsf <- fledging_modelf %>%
  epred_draws(
    newdata = expand.grid(
      TIME = seq(min(dat_female_fledging$TIME), max(dat_female_fledging$TIME), length.out = 100),
      AGE = mean(dat_female_fledging$AGE),
      ZX5MISMATCH = mean(dat_female_fledging$ZX5MISMATCH),
      AFR = mean(dat_female_fledging$AFR),
      LONGEVITY = mean(dat_female_fledging$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_time = TIME + mean_fledging_yearf)

fledging_success_drawsf <- fledging_time_drawsf %>%
  filter(.category != "0") %>%
  group_by(across(-c(.epred, .category))) %>%
  summarize(prob_success = sum(.epred), .groups = "drop")

mean_fledging_yearm <- mean(dat_male_fledging$ori_time)
fledging_time_drawsm <- fledging_modelm %>%
  epred_draws(
    newdata = expand.grid(
      TIME = seq(min(dat_male_fledging$TIME), max(dat_male_fledging$TIME), length.out = 100),
      AGE = mean(dat_male_fledging$AGE),
      ZX5MISMATCH = mean(dat_male_fledging$ZX5MISMATCH),
      AFR = mean(dat_male_fledging$AFR),
      LONGEVITY = mean(dat_male_fledging$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_time = TIME + mean_fledging_yearm)

fledging_success_drawsm <- fledging_time_drawsm %>%
  filter(.category != "0") %>%
  group_by(across(-c(.epred, .category))) %>%
  summarize(prob_success = sum(.epred), .groups = "drop")

saveRDS(bloom_time_draws, file.path(precomputed_draw_dir, "bloom_time_draws.rds"))
saveRDS(mismatch_time_drawsf, file.path(precomputed_draw_dir, "mismatch_time_drawsf.rds"))
saveRDS(mismatch_time_drawsm, file.path(precomputed_draw_dir, "mismatch_time_drawsm.rds"))
saveRDS(fledging_time_drawsf, file.path(precomputed_draw_dir, "fledging_time_drawsf.rds"))
saveRDS(fledging_success_drawsf, file.path(precomputed_draw_dir, "fledging_success_drawsf.rds"))
saveRDS(fledging_time_drawsm, file.path(precomputed_draw_dir, "fledging_time_drawsm.rds"))
saveRDS(fledging_success_drawsm, file.path(precomputed_draw_dir, "fledging_success_drawsm.rds"))
```

```{r}
#| label: totoro-figure3-draw-recipes
#| eval: false
#| code-fold: true
#| code-summary: "Show totoro recipes for Figure 3 posterior draws"

mean_of_original_age_female <- mean(dat_female_mismatch$ori_age)
age_draws_female <- mismatch_modelf %>%
  epred_draws(
    newdata = expand.grid(
      AGE = seq(min(dat_female_mismatch$AGE), max(dat_female_mismatch$AGE), length.out = 100),
      TIME = mean(dat_female_mismatch$TIME),
      AFR = mean(dat_female_mismatch$AFR),
      LONGEVITY = mean(dat_female_mismatch$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_age = AGE + mean_of_original_age_female)

mean_of_original_age_male <- mean(dat_male_mismatch$ori_age)
age_draws_male <- mismatch_modelm %>%
  epred_draws(
    newdata = expand.grid(
      AGE = seq(min(dat_male_mismatch$AGE), max(dat_male_mismatch$AGE), length.out = 100),
      TIME = mean(dat_male_mismatch$TIME),
      AFR = mean(dat_male_mismatch$AFR),
      LONGEVITY = mean(dat_male_mismatch$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_age = AGE + mean_of_original_age_male)

set.seed(42)
rings_to_plot_female <- sample(unique(dat_female_mismatch$RING), 50)
newdata_grid_female <- expand.grid(
  AGE = seq(min(dat_female_mismatch$AGE), max(dat_female_mismatch$AGE), length.out = 100),
  RING = rings_to_plot_female,
  TIME = mean(dat_female_mismatch$TIME),
  AFR = mean(dat_female_mismatch$AFR),
  LONGEVITY = mean(dat_female_mismatch$LONGEVITY)
)
individual_predictions_female <- mismatch_modelf %>%
  add_linpred_draws(newdata = newdata_grid_female, re_formula = ~(1 + AGE | RING)) %>%
  group_by(RING, AGE) %>%
  summarise(.value = median(.linpred), .groups = "drop") %>%
  mutate(actual_age = AGE + mean_of_original_age_female)

set.seed(42)
rings_to_plot_male <- sample(unique(dat_male_mismatch$RING), 50)
newdata_grid_male <- expand.grid(
  AGE = seq(min(dat_male_mismatch$AGE), max(dat_male_mismatch$AGE), length.out = 100),
  RING = rings_to_plot_male,
  TIME = mean(dat_male_mismatch$TIME),
  AFR = mean(dat_male_mismatch$AFR),
  LONGEVITY = mean(dat_male_mismatch$LONGEVITY)
)
individual_predictions_male <- mismatch_modelm %>%
  add_linpred_draws(newdata = newdata_grid_male, re_formula = ~(1 + AGE | RING)) %>%
  group_by(RING, AGE) %>%
  summarise(.value = median(.linpred), .groups = "drop") %>%
  mutate(actual_age = AGE + mean_of_original_age_male)

mean_of_original_fledging_agef <- mean(dat_female_fledging$ori_age)
fledging_age_drawsf <- fledging_modelf %>%
  epred_draws(
    newdata = expand.grid(
      AGE = seq(min(dat_female_fledging$AGE), max(dat_female_fledging$AGE), length.out = 100),
      TIME = mean(dat_female_fledging$TIME),
      ZX5MISMATCH = mean(dat_female_fledging$ZX5MISMATCH),
      AFR = mean(dat_female_fledging$AFR),
      LONGEVITY = mean(dat_female_fledging$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_age = AGE + mean_of_original_fledging_agef)

fledging_success_draws_agef <- fledging_age_drawsf %>%
  filter(.category != "0") %>%
  group_by(across(-c(.epred, .category))) %>%
  summarize(prob_success = sum(.epred), .groups = "drop")

mean_of_original_fledging_agem <- mean(dat_male_fledging$ori_age)
fledging_age_drawsm <- fledging_modelm %>%
  epred_draws(
    newdata = expand.grid(
      AGE = seq(min(dat_male_fledging$AGE), max(dat_male_fledging$AGE), length.out = 100),
      TIME = mean(dat_male_fledging$TIME),
      ZX5MISMATCH = mean(dat_male_fledging$ZX5MISMATCH),
      AFR = mean(dat_male_fledging$AFR),
      LONGEVITY = mean(dat_male_fledging$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_age = AGE + mean_of_original_fledging_agem)

fledging_success_draws_agem <- fledging_age_drawsm %>%
  filter(.category != "0") %>%
  group_by(across(-c(.epred, .category))) %>%
  summarize(prob_success = sum(.epred), .groups = "drop")

mean_of_original_mismatchf <- mean(dat_female_fledging$X5MISMATCH, na.rm = TRUE)
mismatch_drawsf <- fledging_modelf %>%
  epred_draws(
    newdata = expand.grid(
      ZX5MISMATCH = seq(min(dat_female_fledging$ZX5MISMATCH), max(dat_female_fledging$ZX5MISMATCH), length.out = 100),
      TIME = mean(dat_female_fledging$TIME),
      AGE = mean(dat_female_fledging$AGE),
      AFR = mean(dat_female_fledging$AFR),
      LONGEVITY = mean(dat_female_fledging$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_mismatch = ZX5MISMATCH + mean_of_original_mismatchf)

fledging_grouped_drawsf <- mismatch_drawsf %>%
  mutate(fled_group = case_when(
    .category == "0" ~ "0",
    .category %in% c("1", "2") ~ "1-2",
    .category == "3" ~ "3",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(fled_group)) %>%
  group_by(actual_mismatch, .draw, fled_group) %>%
  summarize(prob_val = sum(.epred), .groups = "drop")

mean_of_original_mismatchm <- mean(dat_male_fledging$X5MISMATCH, na.rm = TRUE)
mismatch_drawsm <- fledging_modelm %>%
  epred_draws(
    newdata = expand.grid(
      ZX5MISMATCH = seq(min(dat_male_fledging$ZX5MISMATCH), max(dat_male_fledging$ZX5MISMATCH), length.out = 100),
      TIME = mean(dat_male_fledging$TIME),
      AGE = mean(dat_male_fledging$AGE),
      AFR = mean(dat_male_fledging$AFR),
      LONGEVITY = mean(dat_male_fledging$LONGEVITY)
    ),
    re_formula = NA,
    dpar = TRUE
  ) %>%
  mutate(actual_mismatch = ZX5MISMATCH + mean_of_original_mismatchm)

fledging_grouped_drawsm <- mismatch_drawsm %>%
  mutate(fled_group = case_when(
    .category == "0" ~ "0",
    .category %in% c("1", "2") ~ "1-2",
    .category == "3" ~ "3",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(fled_group)) %>%
  group_by(actual_mismatch, .draw, fled_group) %>%
  summarize(prob_val = sum(.epred), .groups = "drop")

saveRDS(age_draws_female, file.path(precomputed_draw_dir, "age_draws_female.rds"))
saveRDS(age_draws_male, file.path(precomputed_draw_dir, "age_draws_male.rds"))
saveRDS(individual_predictions_female, file.path(precomputed_draw_dir, "individual_predictions_female.rds"))
saveRDS(individual_predictions_male, file.path(precomputed_draw_dir, "individual_predictions_male.rds"))
saveRDS(fledging_age_drawsf, file.path(precomputed_draw_dir, "fledging_age_drawsf.rds"))
saveRDS(fledging_success_draws_agef, file.path(precomputed_draw_dir, "fledging_success_draws_agef.rds"))
saveRDS(fledging_age_drawsm, file.path(precomputed_draw_dir, "fledging_age_drawsm.rds"))
saveRDS(fledging_success_draws_agem, file.path(precomputed_draw_dir, "fledging_success_draws_agem.rds"))
saveRDS(mismatch_drawsf, file.path(precomputed_draw_dir, "mismatch_drawsf.rds"))
saveRDS(fledging_grouped_drawsf, file.path(precomputed_draw_dir, "fledging_grouped_drawsf.rds"))
saveRDS(mismatch_drawsm, file.path(precomputed_draw_dir, "mismatch_drawsm.rds"))
saveRDS(fledging_grouped_drawsm, file.path(precomputed_draw_dir, "fledging_grouped_drawsm.rds"))
```

## Data preprocessing

Raw data were transformed to create the variables used for plotting and model-based prediction.

### Bloom timing data

```{r}
#| label: preprocess-bloom-data
#| eval: false
#| code-fold: true
#| code-summary: "Show bloom preprocessing"

# --- Process Bloom Data ---
bloom_dat <- bloom_dat_raw %>%
  mutate(
    bloom_start_thr5_as_date = as.Date(bloom_start_thr5, format = "%d/%m/%Y"),
    winter_start_date = as.Date(paste0(season_year - 1, "-12-21")),
    delta5 = as.numeric(difftime(bloom_start_thr5_as_date, winter_start_date, units = "days")),
    season_year = as.numeric(season_year),
    TIME = as.numeric(scale(season_year, center = TRUE, scale = FALSE))
  )
```

::: panel-tabset
## Females

```{r}
#| label: preprocess-female-data
#| eval: false
#| code-fold: true
#| code-summary: "Show female preprocessing"



# --- Process Female Mismatch & Fledging Data ---
dat_female_mismatch <- female_dat_raw %>%
  filter(SEX == 1) %>%
  mutate(
    ori_age = as.numeric(AGE),
    ori_time = as.numeric(YEAR),
    TIME = scale(YEAR, center = TRUE, scale = FALSE),
    AGE = scale(as.numeric(AGE), center = TRUE, scale = FALSE),
    AFR = scale(as.numeric(AFR), center = TRUE, scale = FALSE),
    LONGEVITY = scale(as.numeric(LONGEVITY), center = TRUE, scale = FALSE),
    YEAR = as.factor(YEAR),
    RING = as.factor(RING)
  )

dat_female_fledging <- female_dat_raw %>%
  filter(SEX == 1) %>%
  mutate(
    ori_age = as.numeric(AGE),
    ori_time = as.numeric(YEAR),
    TIME = scale(YEAR, center = TRUE, scale = FALSE),
    ZX5MISMATCH = scale(as.numeric(X5MISMATCH), center = TRUE, scale = FALSE),
    AGE = scale(as.numeric(AGE), center = TRUE, scale = FALSE),
    AFR = scale(as.numeric(AFR), center = TRUE, scale = FALSE),
    LONGEVITY = scale(as.numeric(LONGEVITY), center = TRUE, scale = FALSE),
    FLEDS = factor(FLEDS, levels = c(0, 1, 2, 3), ordered = TRUE)
  )
```

## Males

```{r}
#| label: preprocess-male-data
#| eval: false
#| code-fold: true
#| code-summary: "Show male preprocessing"

dat_male_mismatch <- male_dat_raw %>%
  filter(SEX == 0) %>%
  mutate(
    ori_age = as.numeric(AGE),
    ori_time = as.numeric(YEAR),
    TIME = scale(YEAR, center = TRUE, scale = FALSE),
    AGE = scale(as.numeric(AGE), center = TRUE, scale = FALSE),
    AFR = scale(as.numeric(AFR), center = TRUE, scale = FALSE),
    LONGEVITY = scale(as.numeric(LONGEVITY), center = TRUE, scale = FALSE),
    YEAR = as.factor(YEAR),
    RING = as.factor(RING)
  )

dat_male_fledging <- male_dat_raw %>%
  filter(SEX == 0) %>%
  mutate(
    ori_age = as.numeric(AGE),
    ori_time = as.numeric(YEAR),
    TIME = scale(YEAR, center = TRUE, scale = FALSE),
    ZX5MISMATCH = scale(as.numeric(X5MISMATCH), center = TRUE, scale = FALSE),
    AGE = scale(as.numeric(AGE), center = TRUE, scale = FALSE),
    AFR = scale(as.numeric(AFR), center = TRUE, scale = FALSE),
    LONGEVITY = scale(as.numeric(LONGEVITY), center = TRUE, scale = FALSE),
    FLEDS = factor(FLEDS, levels = c(0, 1, 2, 3), ordered = TRUE))
```
:::

# Figure 2: Temporal Trends in Phenology and Fitness

This figure summarizes temporal trends in bloom timing, phenological mismatch, and fledging success.

## Bloom timing predictions

```{r}
#| label: figure2-bloom-predictions
#| eval: false
#| code-fold: true
#| code-summary: "Show bloom prediction code"

# --- Panel A-B: Bloom onset mean and predictability over time ---
bloom_time_draws <- read_precomputed_draw("bloom_time_draws.rds")
```

## Bloom timing panels

```{r}
#| label: figure2-bloom-panels
#| eval: false
#| code-fold: true
#| code-summary: "Show bloom timing panel code"

# Create plot A
plot_A <- ggplot(bloom_dat, aes(x = season_year, y = delta5)) +
geom_point(shape = 21, stroke = 1, fill = "#bcd2b1",
             colour = "#675d76", size = 2, alpha = 0.8) +
  stat_lineribbon(data = bloom_time_draws, aes(x = actual_year, y = .epred), 
                  .width = 0.95, alpha = 0.4, fill = "#6a795a") +
  stat_lineribbon(data = bloom_time_draws, aes(x = actual_year, y = .epred), 
                  .width = 1, alpha = 1, fill = NA)+
  
  my_x_scale  +
  scale_y_continuous(breaks = seq(0,60, 10),minor_breaks=seq(0,60,5),limits = c(0,60),guide = guide_axis(minor.ticks = TRUE)) +
  labs(
    subtitle = "\u03B2 = 0.532, 95% CI = [-0.382, 1.469]",
    y = "Bloom onset (days)",
    x = NULL
  ) +
  custom_theme()
```

```{r}
#| label: figure-2-residuals-bloom
#| message: false
#| warning: false
#| eval: false

model_residuals_bloom <- residuals(bloom_model, summary = TRUE)[, "Estimate"]
dat_with_residuals_bloom <- bloom_dat%>%
    mutate(raw_sigma = abs(model_residuals_bloom)) 

```

```{r}
#| label: figure-2-plot-B
#| message: false
#| warning: false
#| eval: false

plot_B <- ggplot(dat_with_residuals_bloom, aes(x =season_year , y = raw_sigma))+
  geom_point(shape = 21, stroke = 1, fill = "#69666d",
             colour = "#6a795a", size = 2, alpha = 0.8) +
  stat_lineribbon(data = bloom_time_draws, aes(x = actual_year, y = sigma), 
                  .width = 0.95, alpha = 0.4, fill = "#675d76") +
  stat_lineribbon(data = bloom_time_draws, aes(x = actual_year, y = sigma), 
                  linewidth = 1, color= "black", alpha = 1, fill = NA)+

  my_x_scale  +
  scale_y_continuous(breaks = seq(0,50, 10),minor_breaks = seq(0,50,5),limits = c(0,50),guide = guide_axis(minor.ticks = TRUE)) +
  labs(subtitle =  "\u03B2 = 0.114, 95% CI = [0.048, 0.184]",
    y = "Unpredictability (\u03C3)",
    x = NULL
  ) +
  custom_theme()+theme(plot.subtitle = element_text(face="bold"))
```

Similarly, predictions are generated for phenological mismatch (.epred) and its consistency (sigma) over time for the average female/male.

::: panel-tabset
## Females

```{r}
#| label: figure-2-predictions-mismatch-females
#| message: false
#| warning: false
#| eval: false


# --- Panel C: Consistency of synchrony (mismatch variance) ---
mismatch_time_drawsf <- read_precomputed_draw("mismatch_time_drawsf.rds")
```

## Males

```{r}
#| label: figure-2-predictions-mismatch-males
#| message: false
#| warning: false
#| eval: false

mismatch_time_drawsm <- read_precomputed_draw("mismatch_time_drawsm.rds")

```
:::

To visualize the raw data, model residuals are calculated and summarized to show the observed variance per year.

::: panel-tabset
## Females

```{r}
#| label: figure-2-residuals-females
#| message: false
#| warning: false
#| eval: false

 # Calculate residuals for plotting raw variance points
model_residuals_female <- residuals(mismatch_modelf, summary = TRUE)[, "Estimate"]
dat_with_residuals_female <- dat_female_mismatch %>%
    mutate(raw_sigma = abs(model_residuals_female)) 

dat_count_mismatch_yearf <- dat_female_mismatch %>% group_by(ori_time, X5MISMATCH) %>% summarize(n = n(), .groups = 'drop')
dat_count_sigma_year_female <- dat_with_residuals_female %>% group_by(ori_time, raw_sigma) %>% summarize(n = n(), .groups = 'drop')
```

## Males

```{r}
#| label: figure-2-residuals-males
#| message: false
#| warning: false
#| eval: false

model_residuals_male <- residuals(mismatch_modelm, summary = TRUE)[, "Estimate"]
dat_with_residuals_male <- dat_male_mismatch %>%
    mutate(raw_sigma = abs(model_residuals_male)) 

dat_count_mismatch_yearm <- dat_male_mismatch %>% group_by(ori_time, X5MISMATCH) %>% summarize(n = n(), .groups = 'drop')
dat_count_sigma_year_male <- dat_with_residuals_male %>% group_by(ori_time, raw_sigma) %>% summarize(n = n(), .groups = 'drop')
```
:::

The code below builds Plot C (mean mismatch over time) and Plot D (consistency of mismatch over time).

::: panel-tabset
## Females

```{r}
#| label: figure-2-plot-C-D-females
#| message: false
#| warning: false
#| eval: false


# Create plot C
plot_C <- ggplot() +geom_jitter(data = dat_count_mismatch_yearf, 
                                  aes(x = ori_time, y = X5MISMATCH, size = n), width = 0.2, height = 0,
                                  shape = 21, alpha = 0.2,stroke=0.1, fill = "#d2bfb8", colour = "#9096b1")+
  stat_lineribbon(data = mismatch_time_drawsf, aes(x = actual_time, y = .epred), .width = 0.95, alpha = 0.6, fill = "#a0786e") +
  stat_lineribbon(data = mismatch_time_drawsf, aes(x = actual_time, y = .epred), fill = NA, color = "black", linewidth = 1) +
  my_x_scale  +
  scale_y_continuous(breaks = seq(-90,210, 30), limits = c(-90,210),minor_breaks = seq(-90,210,15),guide = guide_axis(minor.ticks = TRUE)) +
  labs(
    subtitle = "\u03B2 = -1.967, 95% CI = [-5.367, 1.399]",
    y = "Phenological mismatch (days)",
    x = NULL
  ) +
  custom_theme() +
  theme(legend.position = "none")


plot_D <- ggplot() +geom_jitter(data = dat_count_sigma_year_female, 
                                  aes(x = ori_time, y = raw_sigma), width = 0.2, height = 0,
                                  shape = 21, alpha = 0.2,stroke=0.1, fill = "#b1ab90", colour = "#b8cbd2",size=1)+
  stat_lineribbon(data = mismatch_time_drawsf, aes(x = actual_time, y = sigma), .width = 0.95, alpha = 0.6, fill = "#cbba62") +
  stat_lineribbon(data = mismatch_time_drawsf, aes(x = actual_time, y = sigma), fill = NA, color = "black", linewidth = 1) +
  my_x_scale  +
  scale_y_continuous(breaks = seq(0,150, 30), limits = c(0,150),guide = guide_axis(minor.ticks = TRUE)) +
  labs(subtitle = "\u03B2 = 0.023, 95% CI = [0.018, 0.028]",
    y = "Inconsistency (\u03C3)",
    x = NULL
  ) +
  custom_theme() +
  theme(legend.position = "none")+theme(plot.subtitle = element_text(face="bold"))
```

## Males

```{r}
#| label: figure-2-plot-C-D-males
#| message: false
#| warning: false
#| eval: false


# Create plot C
plot_Cm <- ggplot() +geom_point(data = dat_count_mismatch_yearm,
 aes(x = ori_time, y = X5MISMATCH, size = n), width = 0.2, height = 0,
                                  shape = 21, alpha = 0.2,stroke=0.1, fill = "#d2bfb8", colour = "#9096b1")+
  stat_lineribbon(data = mismatch_time_drawsm, aes(x = actual_time, y = .epred), .width = 0.95, alpha = 0.6, fill = "#a0786e") +
  stat_lineribbon(data = mismatch_time_drawsm, aes(x = actual_time, y = .epred), fill = NA, color = "black", linewidth = 1) +
  my_x_scale  +
  scale_y_continuous(breaks = seq(-90,210, 30), limits = c(-90,210),minor_breaks = seq(-90,210,15),guide = guide_axis(minor.ticks = TRUE)) +
  labs(
    subtitle = "\u03B2 = -1.833, 95% CI = [-5.137, 1.456]",
    y = "Phenological mismatch (days)",
    x = NULL
  ) +
  custom_theme() +
  theme(legend.position = "none")

# Create plot D
plot_Dm <- ggplot() +geom_jitter(data = dat_count_sigma_year_male, 
                                  aes(x = ori_time, y = raw_sigma), width = 0.2, height = 0,
                                  shape = 21, alpha = 0.2,stroke=0.1, fill = "#b1ab90", colour = "#b8cbd2",size=1)+
  stat_lineribbon(data = mismatch_time_drawsm, aes(x = actual_time, y = sigma), .width = 0.95, alpha = 0.6, fill = "#cbba62") +
  stat_lineribbon(data = mismatch_time_drawsm, aes(x = actual_time, y = sigma), fill = NA, color = "black", linewidth = 1) +
  my_x_scale  +
  scale_y_continuous(breaks = seq(0,150, 30), limits = c(0,150),guide = guide_axis(minor.ticks = TRUE)) +
  labs(subtitle = "\u03B2 = 0.027, 95% CI = [0.023, 0.032]",
    y = "Inconsistency (\u03C3)",
    x = NULL
  ) +
  custom_theme() +
  theme(legend.position = "none")+theme(plot.subtitle = element_text(face="bold"))
```
:::

Next, predictions are generated for fledgling success over time. This includes the probability of each outcome (.epred) and the consistency (disc).

::: panel-tabset
## Females

```{r}
#| label: figure-2-predictions-fledging-females
#| message: false
#| warning: false
#| eval: false

# --- Panel E-F: Fledging success (mean trend) ---
fledging_time_drawsf <- read_precomputed_draw("fledging_time_drawsf.rds")

# Calculate observed proportions for raw points
fledging_props_timef <- dat_female_fledging %>%
  count(ori_time, FLEDS) %>%
  group_by(ori_time) %>%
  mutate(proportion = n / sum(n))



# 1. Load posterior draws for "success" (fledglings > 0)
fledging_success_drawsf <- read_precomputed_draw("fledging_success_drawsf.rds")

# 2. Calculate observed proportions for "success"
fledging_success_props_timef <- dat_female_fledging %>%
  mutate(SUCCESS = ifelse(FLEDS == "0", "Failure (0)", "Success (1+)")) %>%
  count(ori_time, SUCCESS) %>%
  tidyr::complete(ori_time, 
                  SUCCESS = c("Failure (0)", "Success (1+)"), 
                  fill = list(n = 0)) %>%
  group_by(ori_time) %>%
  mutate(
         total_n_year=sum(n),
         proportion = ifelse(total_n_year > 0, n / total_n_year, 0)) %>%
  filter(SUCCESS == "Success (1+)") %>%
  ungroup()

success_label_dataf <- fledging_success_drawsf %>%
  filter(actual_time == min(actual_time) | actual_time == max(actual_time)) %>%
  group_by(actual_time) %>%
  summarise(mean_success_perc = mean(prob_success) * 100, .groups = 'drop')

success_label_dataf

# 3. Percent decline in EXPECTED FLEDGING OUTPUT (last vs first year).
#    "Fledging output" = the expected number of fledglings, i.e. the
#    .epred-weighted sum over outcome categories (0,1,2,3), NOT the
#    probability of success (>0 fledglings) used above. Computed per
#    posterior draw, then summarised with median + 95% credible interval.
#    This is the quantity reported in-text as "54.3% lower" (median decline).
expected_output_drawsf <- fledging_time_drawsf %>%
  filter(actual_time == min(actual_time) | actual_time == max(actual_time)) %>%
  mutate(cat_num = as.numeric(as.character(.category))) %>%
  group_by(.draw, actual_time) %>%
  summarise(expected_fleds = sum(cat_num * .epred), .groups = 'drop')

pct_decline_output_drawsf <- expected_output_drawsf %>%
  mutate(year_pos = ifelse(actual_time == min(actual_time), "first", "last")) %>%
  select(.draw, year_pos, expected_fleds) %>%
  tidyr::pivot_wider(names_from = year_pos, values_from = expected_fleds) %>%
  mutate(pct_decline = (first - last) / first * 100)

pct_decline_output_summaryf <- pct_decline_output_drawsf %>%
  summarise(
    mean_pct_decline = mean(pct_decline),
    median_pct_decline = median(pct_decline),
    lower95 = quantile(pct_decline, 0.025),
    upper95 = quantile(pct_decline, 0.975)
  )

pct_decline_output_summaryf
```

## Males

```{r}
#| label: figure-1-predictions-fledging-males
#| message: false
#| warning: false
#| eval: false

fledging_time_drawsm <- read_precomputed_draw("fledging_time_drawsm.rds")

# Calculate observed proportions for raw points
fledging_props_timem <- dat_male_fledging %>%
  count(ori_time, FLEDS) %>%
  group_by(ori_time) %>%
  mutate(proportion = n / sum(n))


# 1. Load posterior draws for "success" (fledglings > 0)
fledging_success_drawsm <- read_precomputed_draw("fledging_success_drawsm.rds")

# 2. Calculate observed proportions for "success"
fledging_success_props_timem <- dat_male_fledging %>%
  mutate(SUCCESS = ifelse(FLEDS == "0", "Failure (0)", "Success (1+)")) %>%
  count(ori_time, SUCCESS) %>%
  tidyr::complete(ori_time, 
                  SUCCESS = c("Failure (0)", "Success (1+)"), 
                  fill = list(n = 0)) %>%
  group_by(ori_time) %>%
  mutate(
         total_n_year=sum(n),
         proportion = ifelse(total_n_year > 0, n / total_n_year, 0)) %>%
  filter(SUCCESS == "Success (1+)") %>%
  ungroup()


success_label_datam <- fledging_success_drawsm %>%
  filter(actual_time == min(actual_time) | actual_time == max(actual_time)) %>%
  group_by(actual_time) %>%
  summarise(mean_success_perc = mean(prob_success) * 100, .groups = 'drop')
success_label_datam

# 3. Percent decline in EXPECTED FLEDGING OUTPUT (last vs first year).
#    "Fledging output" = the expected number of fledglings, i.e. the
#    .epred-weighted sum over outcome categories (0,1,2,3), NOT the
#    probability of success (>0 fledglings) used above. Computed per
#    posterior draw, then summarised with median + 95% credible interval.
expected_output_drawsm <- fledging_time_drawsm %>%
  filter(actual_time == min(actual_time) | actual_time == max(actual_time)) %>%
  mutate(cat_num = as.numeric(as.character(.category))) %>%
  group_by(.draw, actual_time) %>%
  summarise(expected_fleds = sum(cat_num * .epred), .groups = 'drop')

pct_decline_output_drawsm <- expected_output_drawsm %>%
  mutate(year_pos = ifelse(actual_time == min(actual_time), "first", "last")) %>%
  select(.draw, year_pos, expected_fleds) %>%
  tidyr::pivot_wider(names_from = year_pos, values_from = expected_fleds) %>%
  mutate(pct_decline = (first - last) / first * 100)

pct_decline_output_summarym <- pct_decline_output_drawsm %>%
  summarise(
    mean_pct_decline = mean(pct_decline),
    median_pct_decline = median(pct_decline),
    lower95 = quantile(pct_decline, 0.025),
    upper95 = quantile(pct_decline, 0.975)
  )

pct_decline_output_summarym
```
:::

The next two chunks create Plot E (probability of different fledgling numbers over time) and Plot F (consistency of fledgling success over time).

::: panel-tabset
## Females

```{r}
#| label: figure-2-plot-E-F-females
#| message: false
#| warning: false
#| eval: false
plot_E <- ggplot() +
  geom_point(data = fledging_success_props_timef, 
             aes(x = ori_time, y = proportion,size=total_n_year), 
             shape = 21, alpha = 0.4, stroke=1, fill = "#9dbcbe", colour = "#a68551")+
  # Plot posterior ribbon for success
  stat_lineribbon(data = fledging_success_drawsf, aes(x = actual_time, y = prob_success), 
                  .width = 0.95, alpha = 0.4, fill = "#5172a6") + 
  stat_lineribbon(data = fledging_success_drawsf, aes(x = actual_time, y = prob_success), 
                  fill = NA, color = "black", linewidth = 1) +
  my_x_scale +
  scale_y_continuous(breaks = seq(0, 1, 0.2), minor_breaks = seq(0,1,0.1),limits = c(0, 1),guide = guide_axis(minor.ticks = TRUE)) +
  labs(
    subtitle = "\u03B2 = -0.419, 95% CI = [-1.073, -0.034]",
    y = "Probability of nest success",
    x = NULL
  ) +
  custom_theme()+theme(plot.subtitle = element_text(face="bold"))


plot_F <- ggplot(fledging_time_drawsf, aes(x = actual_time, y = disc)) +
  stat_lineribbon(fill = "#a68551", .width = 0.95, alpha = 0.6, color = NA)  +
  stat_lineribbon(data = fledging_time_drawsf, aes(x = actual_time, y = disc), fill = NA, color = "black", linewidth = 1)+
 my_x_scale  +
  scale_y_continuous(breaks = seq(0,0.5, 0.1), limits = c(0,0.5),guide = guide_axis(minor.ticks = TRUE)) +
  labs(subtitle = "\u03B2 = -0.015, 95% CI = [-0.020, -0.009]", 
       y = "Consistency (\u03B6)",
       x = NULL) +
  custom_theme()+theme(plot.subtitle = element_text(face="bold"))


```

## Males

```{r}
#| label: figure-2-plot-E-F-males
#| message: false
#| warning: false
#| eval: false



plot_Em <- ggplot() +
  geom_point(data = fledging_success_props_timem, 
             aes(x = ori_time, y = proportion, size=total_n_year), 
             shape = 21, alpha = 0.4, stroke=1, fill = "#9dbcbe", colour = "#a68551")+
  # Plot posterior ribbon for success
  stat_lineribbon(data = fledging_success_drawsm, aes(x = actual_time, y = prob_success), 
                  .width = 0.95, alpha = 0.4, fill = "#5172a6") + 
  stat_lineribbon(data = fledging_success_drawsm, aes(x = actual_time, y = prob_success), 
                  fill = NA, color = "black", linewidth = 1) +
  my_x_scale +
  scale_y_continuous(breaks = seq(0, 1, 0.2), minor_breaks = seq(0,1,0.1), limits = c(0, 1),guide = guide_axis(minor.ticks = TRUE)) +
  labs(
    subtitle = "\u03B2 = -0.483, 95% CI = [-1.246, -0.066]",
    y = "Probability of nest success",
    x = NULL
  ) +
  custom_theme()+theme(plot.subtitle = element_text(face="bold"))


plot_Fm <- ggplot(fledging_time_drawsm, aes(x = actual_time, y = disc)) +
  stat_lineribbon(fill = "#a68551", .width = 0.95, alpha = 0.6, color = NA)  +
  stat_lineribbon(data = fledging_time_drawsm, aes(x = actual_time, y = disc), fill = NA, color = "black", linewidth = 1)+
 my_x_scale  +
  scale_y_continuous(breaks = seq(0,0.5, 0.1), limits = c(0,0.5),guide = guide_axis(minor.ticks = TRUE)) +
  labs(subtitle = "\u03B2 = -0.016, 95% CI = [-0.021, -0.011]", 
       y = "Consistency (\u03B6)",
       x = NULL) +
  custom_theme()+theme(plot.subtitle = element_text(face="bold"))









```
:::

Finally, the individual plots (A-F) are assembled into a single multi-panel figure using the patchwork package.

```{r}
#| label: final-figure2
#| fig-width: 7
#| fig-height: 9
#| dpi: 300
#| eval: false
#| message: false
#| warning: false

figure2_female <- 
                (plot_A | plot_B) /
                (plot_C | plot_D) /
                (plot_E  | plot_F) +
  plot_annotation(
    tag_levels = 'A'
  )

# Print the final figure
figure2_female




```

```{r}
#| label: final-figure2_males
#| fig-width: 7
#| fig-height: 9
#| dpi: 300
#| eval: false
#| message: false
#| warning: false

figure2_male <- (plot_Cm | plot_Dm) /
                (plot_Em  | plot_Fm) +
  plot_annotation(
    tag_levels = 'A'
  )

# Print the final figure
figure2_male
```

The assembled figure is then saved as both a PDF and JPG file.

```{r}
#| label: save-figure.
#| eval: false

ggsave(filename = here("..","Plots", "panel_fig2_females_time.pdf"),
       plot = figure2_female,width = 7,  
       height = 9,
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "panel_fig2_females_time.jpg"),
       plot = figure2_female,
       width = 7,  
       height = 9,  
       dpi = 300)




ggsave(filename = here("..","Plots", "panel_fig2_males_time.pdf"),
       plot = figure2_male,width = 7,  
       height = 7, 
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "panel_fig2_males_time.jpg"),
       plot = figure2_male,
       width = 7,  
       height = 7,  
       dpi = 300)


```

# Figure 3: The Effect of Female/males Age

This figure summarizes age-related patterns in phenological mismatch and fledging success, together with the association between mismatch and fledging outcomes.

::: panel-tabset
## Females

```{r}
#| label: generate-age-predictions-females
#| message: false
#| warning: false
#| eval: false

age_draws_female <- read_precomputed_draw("age_draws_female.rds")

dat_count_mismatch_age_female <- dat_female_mismatch %>%
  group_by(ori_age, X5MISMATCH) %>%
  summarize(n = n(), .groups = 'drop')

dat_count_sigma_age_female <- dat_with_residuals_female %>%
  group_by(ori_age, raw_sigma) %>%
  summarize(n = n(), .groups = 'drop')
```

## Males

```{r}
#| label: generate-age-predictions-males
#| message: false
#| warning: false
#| eval: false
age_draws_male <- read_precomputed_draw("age_draws_male.rds")

dat_count_mismatch_age_male <- dat_male_mismatch %>%
  group_by(ori_age, X5MISMATCH) %>%
  summarize(n = n(), .groups = 'drop')


dat_count_sigma_age_male <- dat_with_residuals_male %>%
  group_by(ori_age, raw_sigma) %>%
  summarize(n = n(), .groups = 'drop')
```
:::

To visualize individual variation, predictions for a random sample of 20 individual birds are generated, showing their unique age-related trajectories.

::: panel-tabset
## Females

```{r}
#| label: generate-age-predictions-ind-females
#| message: false
#| warning: false
#| eval: false


individual_predictions_female <- read_precomputed_draw("individual_predictions_female.rds")
```

## Males

```{r}
#| label: generate-age-predictions-ind-males
#| message: false
#| warning: false
#| eval: false

individual_predictions_male <- read_precomputed_draw("individual_predictions_male.rds")
```
:::

The next two chunks create Plot G (mismatch vs. age) and Plot H (consistency vs. age).

::: panel-tabset
## Females

```{r}
#| label: create-age-panels-females
#| message: false
#| warning: false
#| eval: false

# --- Panel G: Mismatch vs. Age ---
plot_G <- ggplot() +
  geom_point(data = dat_count_mismatch_age_female, aes(x = ori_age, y = X5MISMATCH, size = n),
             shape = 21, alpha = 0.2, color = "#9096b1", fill = "#d2bfb8") +
  stat_lineribbon(data = age_draws_female, aes(x = actual_age, y = .epred),
                  .width = 0.95, alpha = 0.4, fill = "#a0786e") +
  #geom_line(data = individual_predictions_female, aes(x = actual_age, y = .value, group = RING),
            #color = "#D95F02", linewidth = 0.5, alpha = 0.6)+
  stat_lineribbon(data = age_draws_female, aes(x = actual_age, y = .epred),
                  fill = NA, color = "black", linewidth = 1) +
  scale_x_continuous(breaks = seq(1, 23, 2),limits = c(1,23),guide = guide_axis(minor.ticks = TRUE)) +
  scale_y_continuous(breaks = seq(-90, 210, 30),minor_breaks = seq(-90,210,15), limits = c(-90, 210),guide = guide_axis(minor.ticks = TRUE)) +
  labs(title = "\u03B2 age = -0.942, 95% CI = [-1.216, -0.666]",
    subtitle = "\u03B2 age<sup>2</sup> = 0.334, 95% CI = [0.291, 0.377]",
    y = "Phenological mismatch (days)",
    x = "Age (years)"
  ) +
  custom_theme() +
  theme(legend.position = "none")+theme(
    plot.title = element_text(size = rel(0.9)),
    plot.subtitle = ggtext::element_markdown(face="bold", size = rel(0.9))
  )

# --- Panel H: Consistency vs. Age ---
plot_H <- ggplot() +
  geom_jitter(data = dat_count_sigma_age_female, aes(x = ori_age, y = raw_sigma), width = 0.2, height = 0, shape = 21, alpha = 0.2,stroke=0.1, fill = "#b1ab90", colour = "#b8cbd2",size=1) +
  stat_lineribbon(data = age_draws_female, aes(x = actual_age, y = sigma),
                  .width = 0.95, alpha = 0.4, fill = "#cbba62") +
  stat_lineribbon(data = age_draws_female, aes(x = actual_age, y = sigma),
                  fill = NA, color = "black", linewidth = 1) +
  scale_x_continuous(breaks = seq(1, 23, 2),limits = c(1,23),guide = guide_axis(minor.ticks = TRUE)) +
  scale_y_continuous(breaks = seq(0, 150, 30), limits = c(0, 150),guide = guide_axis(minor.ticks = TRUE)) +
  labs(title = "\u03B2 age = -0.013, 95% CI = [-0.020, -0.006]",
    subtitle = "\u03B2 age<sup>2</sup> = 0.003, 95% CI = [0.001, 0.004]",
    y = "Inconsistency (\u03C3)",
    x = "Age (years)"
  ) +
  custom_theme() +
  theme(legend.position = "none")+theme(
    plot.title = element_text(size = rel(0.9)),
    plot.subtitle = ggtext::element_markdown(face="bold", size = rel(0.9))
  )
```

## Males

```{r}
#| label: create-age-panels-males
#| message: false
#| warning: false
#| eval: false

# --- Panel G: Mismatch vs. Age ---
plot_Gm <- ggplot() +
  geom_point(data = dat_count_mismatch_age_male, aes(x = ori_age, y = X5MISMATCH, size = n),
             shape = 21, alpha = 0.2, color = "#9096b1", fill = "#d2bfb8") +
  stat_lineribbon(data = age_draws_male, aes(x = actual_age, y = .epred),
                  .width = 0.95, alpha = 0.4, fill = "#a0786e") +
 # geom_line(data = individual_predictions_female, aes(x = actual_age, y = .value, group = RING),
  #          color = "#D95F02", linewidth = 0.5, alpha = 0.6)+
  stat_lineribbon(data = age_draws_male, aes(x = actual_age, y = .epred),
                  fill = NA, color = "black", linewidth = 1) +
  scale_x_continuous(breaks = seq(1, 23, 2),limits = c(1,23),guide = guide_axis(minor.ticks = TRUE)) +
  scale_y_continuous(breaks = seq(-90, 210, 30),minor_breaks = seq(-90,210,15), limits = c(-90, 210),guide = guide_axis(minor.ticks = TRUE)) +
  labs(title = "\u03B2 age = -1.096, 95% CI = [-1.369, -0.823]",
    subtitle = "\u03B2 age<sup>2</sup> = 0.283, 95% CI = [0.235, 0.332]",
    y = "Phenological mismatch (days)",
    x = "Age (years)"
  ) +
  custom_theme() +
  theme(legend.position = "none")+theme(
    plot.title = element_text(size = rel(0.9)),
    plot.subtitle = ggtext::element_markdown(face="bold", size = rel(0.9))
  )

# --- Panel H: Consistency vs. Age ---
plot_Hm <- ggplot() +
  geom_jitter(data = dat_count_sigma_age_male, aes(x = ori_age, y = raw_sigma), width = 0.2, height = 0, shape = 21, alpha = 0.2,stroke=0.1, fill = "#b1ab90", colour = "#b8cbd2",size=1) +
  stat_lineribbon(data = age_draws_male, aes(x = actual_age, y = sigma),
                  .width = 0.95, alpha = 0.4, fill = "#cbba62") +
  stat_lineribbon(data = age_draws_male, aes(x = actual_age, y = sigma),
                  fill = NA, color = "black", linewidth = 1) +
  scale_x_continuous(breaks = seq(1, 23, 2),limits = c(1,23),guide = guide_axis(minor.ticks = TRUE)) +
  scale_y_continuous(breaks = seq(0, 150, 30), limits = c(0, 150),guide = guide_axis(minor.ticks = TRUE)) +
  labs(title = "\u03B2 age = -0.012, 95% CI = [-0.019, -0.005]",
    subtitle = "\u03B2 age<sup>2</sup> = 0.004, 95% CI = [0.003, 0.005]",
    y = "Inconsistency (\u03C3)",
    x = "Age (years)"
  ) +
  custom_theme() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = rel(0.9)),
    plot.subtitle = ggtext::element_markdown(face = "bold", size = rel(0.9))
  )
```
:::

Predictions are now generated from the fledgling success model across the range of female ages.

::: panel-tabset
## Females

```{r}
#| label: generate-fledging-age-predictions-females
#| message: false
#| warning: false
#| eval: false

fledging_age_drawsf <- read_precomputed_draw("fledging_age_drawsf.rds")
fledging_success_draws_agef <- read_precomputed_draw("fledging_success_draws_agef.rds")

fledging_success_props_agef <- dat_female_fledging%>%
  mutate(SUCCESS = ifelse(FLEDS == "0", "Failure (0)", "Success (1+)")) %>%
  count(ori_age, SUCCESS) %>%
  tidyr::complete(ori_age, 
                  SUCCESS = c("Failure (0)", "Success (1+)"), 
                  fill = list(n = 0)) %>%
  group_by(ori_age) %>%
  mutate(
         total_n_year = sum(n),
         proportion = ifelse(total_n_year > 0, n / total_n_year, 0)) %>%
  filter(SUCCESS == "Success (1+)") %>%
  ungroup()



```

## Males

```{r}
#| label: generate-fledging-age-predictions-males
#| message: false
#| warning: false
#| eval: false

fledging_age_drawsm <- read_precomputed_draw("fledging_age_drawsm.rds")
fledging_success_draws_agem <- read_precomputed_draw("fledging_success_draws_agem.rds")

fledging_success_props_agem <- dat_male_fledging%>%
  mutate(SUCCESS = ifelse(FLEDS == "0", "Failure (0)", "Success (1+)")) %>%
  count(ori_age, SUCCESS) %>%
  tidyr::complete(ori_age, 
                  SUCCESS = c("Failure (0)", "Success (1+)"), 
                  fill = list(n = 0)) %>%
  group_by(ori_age) %>%
  mutate(
         total_n_year = sum(n),
         proportion = ifelse(total_n_year > 0, n / total_n_year, 0)) %>%
  filter(SUCCESS == "Success (1+)") %>%
  ungroup()

```
:::

The next chunks build the bottom row of the figure: fledgling probabilities vs. age.

::: panel-tabset
## Females

```{r}
#| label: create-fledging-age-panels-females
#| message: false
#| warning: false
#| eval: false

# --- PanelA: Fledging Probability vs. Age ---
plot_fledging_success_age <- ggplot() +

  geom_point(data = fledging_success_props_agef, 
             aes(x = ori_age, y = proportion,size=total_n_year), 
             shape = 21, alpha = 0.4, stroke = 1, fill = "#9dbcbe", colour = "#a68551") +
  stat_lineribbon(data = fledging_success_draws_agef, 
                  aes(x = actual_age, y = prob_success), 
                  .width = 0.95, alpha = 0.4, fill = "#5172a6") + 
  stat_lineribbon(data = fledging_success_draws_agef, 
                  aes(x = actual_age, y = prob_success), 
                  fill = NA, color = "black", linewidth = 1) +
  

  scale_x_continuous(
    breaks = seq(1, 23, 2),
    minor_breaks = seq(1, 23, 1),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_y_continuous(
    breaks = seq(0, 1, 0.2), 
    minor_breaks = seq(0, 1, 0.1),
    limits = c(0, 1),
    guide = guide_axis(minor.ticks = TRUE)
  ) +

  labs(title = "\u03B2 age = 0.366, 95% CI = [0.150, 0.759]",
    subtitle = "\u03B2 age<sup>2</sup> = -0.075, 95% CI = [-0.156, -0.031]",
    y = "Probability of nest success",
    x = "Age (years)"
  ) +

  custom_theme(show_legend = FALSE)+theme(
    plot.title = element_text(size = rel(0.9)),
    plot.subtitle = ggtext::element_markdown(face="bold", size = rel(0.9))
  )


```

## Males

```{r}
#| label: create-fledging-age-panels-males
#| message: false
#| warning: false
#| eval: false

plot_fledging_success_agem <- ggplot() +

  geom_point(data = fledging_success_props_agem, 
             aes(x = ori_age, y = proportion, size=total_n_year), 
             shape = 21, alpha = 0.4, stroke = 1, fill = "#9dbcbe", colour = "#a68551") +
  stat_lineribbon(data = fledging_success_draws_agem, 
                  aes(x = actual_age, y = prob_success), 
                  .width = 0.95, alpha = 0.4, fill = "#5172a6") + 
  stat_lineribbon(data = fledging_success_draws_agem, 
                  aes(x = actual_age, y = prob_success), 
                  fill = NA, color = "black", linewidth = 1) +
  

  scale_x_continuous(
    breaks = seq(1, 23, 2),
    minor_breaks = seq(1, 23, 1),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_y_continuous(
    breaks = seq(0, 1, 0.2), 
    minor_breaks = seq(0, 1, 0.1),
    limits = c(0, 1),
    guide = guide_axis(minor.ticks = TRUE)
  ) +

  labs(title = "\u03B2 age = 0.171, 95% CI = [0.054, 0.401]",
    subtitle = "\u03B2 age<sup>2</sup> = -0.039, 95% CI = [-0.088, -0.014]",
    y = "Probability of nest success",
    x = "Age (years)"
  ) +


  custom_theme(show_legend = FALSE)+theme(
    plot.title = element_text(size = rel(0.9)),
    plot.subtitle = ggtext::element_markdown(face="bold", size = rel(0.9))
  )

```
:::

This panel shows the direct relationship between phenological mismatch and the probability of fledgling success.

Model predictions are generated for fledgling success across the full range of observed mismatch values.

::: panel-tabset
## Females

```{r}
#| label: generate-fledging-mismatch-predictions-females
#| message: false
#| warning: false
#| eval: false

# --- 1. Load raw and grouped predictions ---
mismatch_drawsf <- read_precomputed_draw("mismatch_drawsf.rds")
fledging_grouped_drawsf <- read_precomputed_draw("fledging_grouped_drawsf.rds")

# --- 3. Calculate Observed Proportions for the 3 Categories ---
fledging_grouped_propsf <- dat_female_fledging %>%
  filter(!is.na(X5MISMATCH)) %>%
  # Create the same grouping variable in the raw data
  mutate(
    fled_group = case_when(
      FLEDS == "0" ~ "0",
      FLEDS %in% c("1", "2") ~ "1-2",
      FLEDS == "3" ~ "3",
      TRUE ~ "Other"
    )
  ) %>%
  filter(fled_group %in% c("0", "1-2", "3")) %>%
  # Count occurrences per mismatch value and group
  count(X5MISMATCH, fled_group) %>%
  # Ensure all groups exist for all mismatch values (fill with 0 if missing)
  tidyr::complete(X5MISMATCH, 
                  fled_group = c("0", "1-2", "3"), 
                  fill = list(n = 0)) %>%
  # Calculate proportions
  group_by(X5MISMATCH) %>%
  mutate(total_n_year = sum(n),
         proportion = ifelse(total_n_year > 0, n / total_n_year, 0)) %>%
  ungroup()

# --- 4. Generate Label Data for Plotting ---
group_mismatch_label_dataf <- fledging_grouped_drawsf %>%
  filter(actual_mismatch == min(actual_mismatch) | actual_mismatch == max(actual_mismatch)) %>%
  group_by(actual_mismatch, fled_group) %>%
  summarise(mean_prob_perc = mean(prob_val) * 100, .groups = 'drop')

group_mismatch_label_dataf
```

## Males

```{r}
#| label: generate-fledging-mismatch-predictions-m
#| message: false
#| warning: false
#| eval: false

mismatch_drawsm <- read_precomputed_draw("mismatch_drawsm.rds")
fledging_grouped_drawsm <- read_precomputed_draw("fledging_grouped_drawsm.rds")

# --- 3. Calculate Observed Proportions for the 3 Categories ---
fledging_grouped_propsm <- dat_male_fledging %>%
  filter(!is.na(X5MISMATCH)) %>%
  # Create the same grouping variable in the raw data
  mutate(
    fled_group = case_when(
      FLEDS == "0" ~ "0",
      FLEDS %in% c("1", "2") ~ "1-2",
      FLEDS == "3" ~ "3",
      TRUE ~ "Other"
    )
  ) %>%
  filter(fled_group %in% c("0", "1-2", "3")) %>%
  # Count occurrences per mismatch value and group
  count(X5MISMATCH, fled_group) %>%
  # Ensure all groups exist for all mismatch values (fill with 0 if missing)
  tidyr::complete(X5MISMATCH, 
                  fled_group = c("0", "1-2", "3"), 
                  fill = list(n = 0)) %>%
  # Calculate proportions
  group_by(X5MISMATCH) %>%
  mutate(total_n_year = sum(n),
         proportion = ifelse(total_n_year > 0, n / total_n_year, 0)) %>%
  ungroup()

# --- 4. Generate Label Data for Plotting ---
group_mismatch_label_datam <- fledging_grouped_drawsm %>%
  filter(actual_mismatch == min(actual_mismatch) | actual_mismatch == max(actual_mismatch)) %>%
  group_by(actual_mismatch, fled_group) %>%
  summarise(mean_prob_perc = mean(prob_val) * 100, .groups = 'drop')

group_mismatch_label_datam
```
:::

::: panel-tabset
## Females

```{r}
#| label: create-fledging-mismatch-panels-females
#| message: false
#| warning: false
#| eval: false


plot_mismatch_prob_altf <- ggplot() + 
  # 1. Observed Data Points (binned proportions)
  # Mapped fill to fled_group to see observed 0s vs 1-2s vs 3s
  geom_point(data = fledging_grouped_propsf, 
             aes(x = X5MISMATCH, y = proportion, size = total_n_year, fill = fled_group), 
             shape = 21, alpha = 0.2, stroke = 0.5, color = "black") +

  # 2. Posterior Ribbons (95% CI)
  stat_lineribbon(data = fledging_grouped_drawsf, 
                  aes(x = actual_mismatch, y = prob_val, fill = fled_group), 
                  .width = 0.95, alpha = 0.3, color = NA) +

  # 3. Posterior Mean Lines
  # We use .width = 0 to draw just the line without another ribbon on top
  stat_lineribbon(data = fledging_grouped_drawsf, 
                  aes(x = actual_mismatch, y = prob_val, color = fled_group), 
                  .width = 0, fill = NA, linewidth = 1) +

  # 4. Scales and Colors
  scale_y_continuous(
    breaks = seq(0, 1, 0.2), 
    minor_breaks = seq(0, 1, 0.1),  
    limits = c(0, 1),
    guide = guide_axis(minor.ticks = TRUE) 
  ) +
  scale_x_continuous(
    breaks = seq(-90, 210, by = 30), 
    minor_breaks = seq(-90, 210, by = 15), 
    limits = c(-90, 210),
    guide = guide_axis(minor.ticks = TRUE) 
  ) +
  
  # Manual colors: Adjust these hex codes to your preference
  scale_fill_manual(values = c("0" = "#999999", "1-2" = "#E69F00", "3" = "#56B4E9"), name = "Number of fledglings") +
  scale_color_manual(values = c("0" = "#999999", "1-2" = "#E69F00", "3" = "#56B4E9"), name = "Number of fledglings") +

  # 5. Labels
  labs(title = "\u03B2 mismatch = -0.184, 95% CI = [-0.374, -0.078]",
    subtitle = "\u03B2 mismatch<sup>2</sup> = -0.0003, 95% CI = [-0.0007, -0.0001]",
    y = "Probability of fledging outcomes",
    x = "Phenological mismatch (days)"
  ) +
  guides(size = "none",
    # title.position = "left" puts the title next to the keys
    # nrow = 1 forces all keys to be in one line
    fill = guide_legend(title.position = "left", nrow = 1),
    color = guide_legend(title.position = "left", nrow = 1))+
  custom_theme() + theme(
    plot.title = element_text(size = rel(0.9)),
    plot.subtitle = ggtext::element_markdown(face = "bold", color = "black", size = rel(0.9)),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),

    # Move legend to bottom
    legend.position = "bottom",
    # Ensure items are not too spaced out
    legend.spacing.x = unit(0.5, "cm")
  )
```

## Males

```{r}
#| label: create-fledging-mismatch-panels-males
#| message: false
#| warning: false
#| eval: false
plot_mismatch_prob_altm <- ggplot() + 
  # 1. Observed Data Points (binned proportions)
  # Mapped fill to fled_group to see observed 0s vs 1-2s vs 3s
  geom_point(data = fledging_grouped_propsm, 
             aes(x = X5MISMATCH, y = proportion, size = total_n_year, fill = fled_group), 
             shape = 21, alpha = 0.2, stroke = 0.5, color = "black") +

  # 2. Posterior Ribbons (95% CI)
  stat_lineribbon(data = fledging_grouped_drawsm, 
                  aes(x = actual_mismatch, y = prob_val, fill = fled_group), 
                  .width = 0.95, alpha = 0.3, color = NA) +

  # 3. Posterior Mean Lines
  # We use .width = 0 to draw just the line without another ribbon on top
  stat_lineribbon(data = fledging_grouped_drawsm, 
                  aes(x = actual_mismatch, y = prob_val, color = fled_group), 
                  .width = 0, fill = NA, linewidth = 1) +

  # 4. Scales and Colors
  scale_y_continuous(
    breaks = seq(0, 1, 0.2), 
    minor_breaks = seq(0, 1, 0.1),  
    limits = c(0, 1),
    guide = guide_axis(minor.ticks = TRUE) 
  ) +
  scale_x_continuous(
    breaks = seq(-90, 210, by = 30), 
    minor_breaks = seq(-90, 210, by = 15), 
    limits = c(-90, 210),
    guide = guide_axis(minor.ticks = TRUE) 
  ) +
  
  # Manual colors: Adjust these hex codes to your preference
  scale_fill_manual(values = c("0" = "#999999", "1-2" = "#E69F00", "3" = "#56B4E9"), name = "Number of fledglings") +
  scale_color_manual(values = c("0" = "#999999", "1-2" = "#E69F00", "3" = "#56B4E9"), name = "Number of fledglings") +

  # 5. Labels
  labs(title = "\u03B2 mismatch = -0.193, 95% CI = [-0.405, -0.077]",
    subtitle = "\u03B2 mismatch<sup>2</sup> = -0.0002, 95% CI = [-0.0005, -0.0000]",
    y = "Probability of fledging outcomes",
    x = "Phenological mismatch (days)"
  ) +
  guides(size = "none",
    # title.position = "left" puts the title next to the keys
    # nrow = 1 forces all keys to be in one line
    fill = guide_legend(title.position = "left", nrow = 1),
    color = guide_legend(title.position = "left", nrow = 1))+
  custom_theme() + theme(
    plot.title = element_text(size = rel(0.9)),
    plot.subtitle = ggtext::element_markdown(face = "bold", color = "black", size = rel(0.9)),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),

    # Move legend to bottom
    legend.position = "bottom",
    # Ensure items are not too spaced out
    legend.spacing.x = unit(0.5, "cm")
)
```
:::

```{r}
#| label: figure 3 females
#| fig-width: 7
#| fig-height: 7
#| dpi: 300
#| eval: false
#| message: false
#| warning: false

figure3_female <- (plot_G | plot_H) /
                (plot_fledging_success_age  | plot_mismatch_prob_altf ) +
  plot_annotation(
    tag_levels = 'A'
  )

# Print the final figure
figure3_female
```

```{r}
#| label: figure 3 
#| fig-width: 7
#| fig-height: 7
#| dpi: 300
#| eval: false
#| message: false
#| warning: false

figure3_male <- (plot_Gm | plot_Hm) /
                (plot_fledging_success_agem  | plot_mismatch_prob_altm ) +
  plot_annotation(
    tag_levels = 'A'
  )

# Print the final figure
figure3_male
```

The assembled figure is then saved as both a PDF and JPG file.

```{r}
#| label: save-figure
#| eval: false

ggsave(filename = here("..","Plots", "panel_fig3_females.pdf"),
       plot = figure3_female, width = 7,  
       height = 9, 
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "panel_fig3_females.jpg"),
       plot = figure3_female,
       width = 7,  
       height = 9,  
       dpi = 300)




ggsave(filename = here("..","Plots", "panel_fig3_males.pdf"),
       plot = figure3_male,
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "panel_fig3_males.jpg"),
       plot = figure3_male, 
       width = 7,  
       height = 9,  
       dpi = 300)


```

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
