---
title: "R Notebook"
output: html_notebook
---

```{r}
#| label: setup-final-summaries
#| include: false

# Load libraries needed for this chapter
pacman::p_load(here, brms, clipr, broom.mixed)
```

::: {.callout-note title="Variable definition"}
\-`X5MISMATCH`: Phenological mismatch calculated using the 5% bloom threshold (in days).

\-`CLUTCH_SIZE`: The number of eggs laid in a nest (ordinal).

\-`FLEDS`: The number of chicks that successfully fledged (ordinal).

\-`TIME`: Chronological time, represented as centered years to model temporal trends. Centering is done by subtracting the mean.

\-`ZX5MISMATCH`: The centered phenological mismatch (5% threshold).

\-`AGE`: The age of the focal individual (in years), centered.

\-`AFR`: The age at first reproduction (in years), centered.

\-`LONGEVITY`: The age at the last observation, a proxy for lifespan (in years), centered.

\-`YEAR`: The calendar year of the observation (treated as a factor).

\-`RING`: The unique alphanumeric identifier for each individual (treated as a factor).
:::

This section presents the posterior summaries for the best-supported models from each primary analysis. The models for Phenological mismatch and Number of fledglings were selected based on Leave-One-Out Cross-Validation (LOO-CV)
```{r}
library(dplyr)
library(tibble)
library(clipr)

copy_summary <- function(model) {
  
  # 1. Get the summary structure
  s <- summary(model)
  
  # 2. Extract Fixed Effects
  # We use tryCatch or simple checks to ensure it doesn't fail if a part is missing
  fixed_df <- if (!is.null(s$fixed)) {
    as.data.frame(s$fixed) %>% 
      rownames_to_column("Term") %>% 
      mutate(Type = "Fixed Effect")
  } else { NULL }
  
  # 3. Extract Random Effects (SDs and Correlations)
  random_df <- if (length(s$random) > 0) {
    # Combine list elements into one matrix, then force to data.frame
    random_raw <- do.call(rbind, lapply(s$random, as.data.frame))
    as.data.frame(random_raw) %>% 
      rownames_to_column("Term") %>% 
      mutate(Type = "Random Effect (SD/Cor)")
  } else { NULL }
  
  # 4. Extract Family Specific Params (Sigma/Shape/etc)
  spec_df <- if (!is.null(s$spec_pars) && nrow(s$spec_pars) > 0) {
    as.data.frame(s$spec_pars) %>% 
      rownames_to_column("Term") %>% 
      mutate(Type = "Family Specific")
  } else { NULL }
  
  # 5. Combine everything
  final_table <- bind_rows(fixed_df, random_df, spec_df)
  
  # 6. Copy to clipboard
  write_clip(final_table)
  
  # Optional: Print a confirmation message to console
  message("✅ Summary copied to clipboard! Ready to paste into Excel.")
  
  # Return the table invisibly in case you want to assign it to a variable
  invisible(final_table)
}
```

```{r}
#| label: summary-bloom
#| eval: true
#| size: "footnotesize"


# Load the model
model_bloom <- readRDS(here::here("..","Rdata", "model_delta5.rds"))

copy_summary(model_bloom)
```
```{r}
library(dplyr)
library(tibble)
library(clipr)

# 1. Capture the summary object
s <- summary(model_bloom)

# 2. Fixed Effects
fixed_df <- as.data.frame(s$fixed) %>% 
  rownames_to_column("Term") %>% 
  mutate(Type = "Fixed Effect")

# 3. Random Effects (The Fix)
# We force the result of do.call into a data.frame immediately
random_raw <- do.call(rbind, lapply(s$random, as.data.frame))
random_df  <- as.data.frame(random_raw) %>% 
  rownames_to_column("Term") %>% 
  mutate(Type = "Random Effect (SD/Cor)")

# 4. Family Specific (Sigma/Residuals)
spec_df <- as.data.frame(s$spec_pars) %>% 
  rownames_to_column("Term") %>% 
  mutate(Type = "Family Specific")

# 5. Bind and Copy
final_table <- bind_rows(fixed_df, random_df, spec_df)

# Write to clipboard for Excel
write_clip(final_table)

```

::::: panel-tabset
## Females

::: panel-tabset
## Phenological Mismatch

```{r}
#| label: summary-pm-female-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_pm_female_best <- readRDS(here::here("..","Rdata", "female_fit_m8.rds"))

# Print the full model summary
# 1. Capture the summary object
s <- summary(model_pm_female_best)

# 2. Fixed Effects
fixed_df <- as.data.frame(s$fixed) %>% 
  rownames_to_column("Term") %>% 
  mutate(Type = "Fixed Effect")

# 3. Random Effects (The Fix)
# We force the result of do.call into a data.frame immediately
random_raw <- do.call(rbind, lapply(s$random, as.data.frame))
random_df  <- as.data.frame(random_raw) %>% 
  rownames_to_column("Term") %>% 
  mutate(Type = "Random Effect (SD/Cor)")

# 4. Family Specific (Sigma/Residuals)
spec_df <- as.data.frame(s$spec_pars) %>% 
  rownames_to_column("Term") %>% 
  mutate(Type = "Family Specific")

# 5. Bind and Copy
final_table <- bind_rows(fixed_df, random_df, spec_df)

# Write to clipboard for Excel
write_clip(final_table)

```

## Number of Fledglings

```{r}
#| label: summary-fs-female-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_fs_female_best <- readRDS(here::here("..","Rdata", "female_fit_Q17.rds"))

# Print the full model summary
print(summary(model_fs_female_best), digits =4)
```
:::

## Males

::: panel-tabset
## Phenological Mismatch

```{r}
#| label: summary-pm-male-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_pm_male_best <- readRDS(here::here("..","Rdata", "male_fit_m8.rds"))

# Print the full model summary
summary(model_pm_male_best)
```

## Number of Fledglings

```{r}
#| label: summary-fs-male-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_fs_male_best <- readRDS(here::here("..","Rdata", "male_fit_Q17.rds"))

# Print the full model summary
print(summary(model_fs_male_best),digits=4)
```