---
title: "Sensitivity analyses: Bloom onset thresholds"
unnumbered: true
---
This section evaluates whether inference about bloom phenology is robust to the chlorophyll-a bloom onset threshold used to define bloom start (5%, 15%, and 25% above the seasonal median). We report posterior summaries for the location–scale models fit to each threshold and provide the code used to fit these models.


```{r}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
    brms, dplyr, here, future, purrr, furrr,cmdstanr,tibble, stringr
)
```

# Threshold summaries

::: panel-tabset
## 5% Threshold

```{r}
#| label: summary-5pct
#| eval: true
summary(readRDS(here("..","Rdata","model_delta5.rds")))
```

## 15% Threshold

```{r}
#| label: summary-15pct
#| eval: true
summary(readRDS(here("..","Rdata","model_delta15.rds")))
```

## 25% Threshold

```{r}
#| label: summary-25pct
#| eval: true
summary(readRDS(here("..","Rdata","model_delta25.rds")))
```
:::

# Model fitting

```{r}
#| label: setup
#| eval: false
# --- Parallel Processing Setup ---
options(mc.cores = parallel::detectCores(), brms.backend = "cmdstanr")
plan(multisession, workers=4) 
```

## Data transformation: bloom onset as days since winter solstice

Bloom onset dates were converted to numeric offsets (days) from the previous winter solstice (December 21), matching the parameterization used in the main bloom phenology analyses.

```{r}

#| label: data-transformation
#| eval: false
dat <- read.csv(here("..","data","bloom_dates.csv"))

dat <- dat %>%
  mutate(
    # --- Date parsing for detected bloom onsets ---
    bloom_start_thr5_as_date  = as.Date(bloom_start_thr5,  format = "%d/%m/%Y"),
    bloom_start_thr15_as_date = as.Date(bloom_start_thr15, format = "%d/%m/%Y"),
    bloom_start_thr25_as_date = as.Date(bloom_start_thr25, format = "%d/%m/%Y"),

    # --- Reference date: previous winter solstice (Dec 21) ---
    winter_start_date = as.Date(paste0(season_year - 1, "-12-21")),

    # --- Bloom timing response variables (days since winter solstice) ---
    delta5  = as.numeric(difftime(bloom_start_thr5_as_date,  winter_start_date, units = "days")),
    delta15 = as.numeric(difftime(bloom_start_thr15_as_date, winter_start_date, units = "days")),
    delta25 = as.numeric(difftime(bloom_start_thr25_as_date, winter_start_date, units = "days")),

    # --- Year variables with distinct roles ---
    # For summaries / tables / bookkeeping only (not used as the model time trend)
    season_year_f = as.factor(season_year),

    # For location–scale models (mean and sigma trends): centered numeric year
    TIME = as.numeric(scale(season_year, center = TRUE, scale = FALSE)),

    # Optional: chlorophyll values at detected bloom onset (kept as numeric)
    value5  = as.numeric(value_at_bloom_thr5),
    value15 = as.numeric(value_at_bloom_thr15),
    value25 = as.numeric(value_at_bloom_thr25)
  ) %>%
  # Drop original date strings and helper date column after transformation
  select(-starts_with("bloom_start_thr"), -winter_start_date)
```
## Fit Gaussian location–scale models

The same Gaussian location–scale model structure was fit to each bloom onset metric (delta5, delta15, delta25), with TIME included in both the location (mean) and scale (residual SD) components.

```{r}
#| label: fit-location-scale-models
#| eval: false
run_location_scale_models <- function(response_vars, data) {
  if (!dir.exists(here("Rdata"))) {
    dir.create(here("Rdata"))
  }
  
  # Use purrr::walk to iterate through each response variable and fit a model.
  purrr::walk(response_vars, function(var) {
    
    # Define the model formula. We model both the mean (mu) and the standard
    # deviation (sigma) as a function of TIME.
    model_formula <- bf(
      as.formula(paste(var, "~ 1 + TIME")),
      sigma ~ 1 + TIME
    )
    
    # Define the output file path for the saved model object.
    file_path <- here("Rdata", paste0("model_", var, ".rds"))
    
    # Define the model family.
    model_family <- gaussian()
    
    # Get the default priors for the specified model formula and data.
    model_priors <- get_prior(model_formula, data = data, family = model_family)
    
    # Use tryCatch to gracefully handle any potential errors during model fitting.
    fit <- tryCatch({
      brm(
        formula = model_formula,
        data = data,
        prior = model_priors,
        family = model_family,
        iter = 6000,
        warmup = 2000,
        chains = 4,
        seed = 123,
        init = 0,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        file = file_path,
        file_refit = "on_change"
      )
    }, error = function(e) {
      # If an error occurs, print a message and return NULL.
      cat("Error fitting model for:", var, "\n")
      cat("Error message:", e$message, "\n")
      return(NULL)
    })
    
    # Only print the success message if the model fitting was successful.
    if (!is.null(fit)) {
      cat("Saved model for:", var, "to", file_path, "\n")
    }
  })
}
```

```{r}
#| label: run-models
#| eval: false
# --- Run Analysis for Delta Variables ---
# Define the list of delta response variables.
delta_vars <- c("delta5", "delta15", "delta25")
# Run the models for the delta variables.
run_location_scale_models(response_vars = delta_vars, data = dat)
```

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
